---
output:
  html_document:
    toc: false
params:
  project_directory: "/Users/isaac/Dropbox/S2P Datasets Shared Folder/MLTC_MMC_NH_data/"
  use_local_data: FALSE
---
<style>
.button_download {
font-size: 16px;
font-family: "Arial";
color: black;
background-color: #C7C9D8;
}
</style>

```{r setup_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, error = TRUE, fig.width = 12, fig.height = 8, cache = FALSE)
```

```{r setup_libraries, include=FALSE}
library(tidyverse)
library(furrr)
library(scales)
library(gt)
library(pdftools)
library(httr)
library(downloadthis)
plan(multisession)
data_dir <- params$project_directory
```

```{r isaac_step2_graph_theme, include=FALSE}
isaac_step2_graph_theme <- theme(plot.title.position = "plot",
                                 plot.title = element_text(face = "bold"),
                                 legend.position = "top",
                                 strip.text = element_text(color = "black"),
                                 legend.title = element_text(face = "bold",hjust = 0.5),
                                 axis.title.y = element_text(face = "bold",angle = 90),
                                 axis.title.x = element_text(face = "bold"))
```

```{r setup_medicaid_enrollment_data_ingestion, include=FALSE}
if(params$use_local_data == FALSE){
  # # https://health.data.ny.gov/Health/Medicaid-Program-Enrollment-by-Month-Beginning-200/m4hz-kzn3/about_data
  # nys_medicaid <- read_csv("https://health.data.ny.gov/api/views/m4hz-kzn3/rows.csv?accessType=DOWNLOAD") %>% janitor::clean_names()
  
  # https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/
  # https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/archives.htm
  
  # Custom headers for the request
  custom_headers <- add_headers(
    `User-Agent` = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/97.0.4692.71 Safari/537.36",
    `Accept` = "application/vnd.openxmlformats-officedocument.spreadsheetml.sheet, application/vnd.ms-excel;q=0.9,*/*;q=0.8",
    `Connection` = "keep-alive"
  )
  
  # Function to download a file
  download_file <- function(url, save_directory, filename) {
    destfile <- file.path(save_directory, filename)  # Create full path
    response <- GET(url, custom_headers)
    
    if (status_code(response) == 200) {
      writeBin(content(response, "raw"), destfile)
      message("File saved to: ", destfile)
    } else {
      warning("Failed to download file: ", url, " - Status code: ", status_code(response))
    }
  }
  
  # Define months and years
  months <- sprintf("%02d", 1:12)  # "01", "02", ..., "12"
  years <- 2008:as.numeric(format(Sys.Date(),"%Y"))              # Adjust range as needed
  
  already_downloaded_mmc_files <- list.files(paste0(data_dir,"mmc_mltc/data"),recursive = TRUE)
  
  # Loop through years and months to construct URLs and download files
  for (year in years) {
    for (month in months) {
      
      if(paste0("en",month,"_",year,".xls") %in% already_downloaded_mmc_files | paste0("en",month,"_",year,".xlsx") %in% already_downloaded_mmc_files){
        message(paste0("en",month,"_",year," was already downloaded"))
      } else{
        
        # Construct the URL for both file types
        url_xlsx <- paste0("https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/", year, "/docs/en", month, "_", substr(year, 3, 4), ".xlsx")
        url_xls <- paste0("https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/", year, "/docs/en", month, "_", substr(year, 3, 4), ".xls")
        
        # Attempt to download .xlsx file first
        tryCatch(
          download_file(url_xlsx, paste0(data_dir,"mmc_mltc/data"), paste0("en", month, "_", year, ".xlsx")),
          error = function(e) {
            message("Trying .xls for ", month, "/", year)
          }
        )
        
        # Attempt to download .xls file if .xlsx fails
        tryCatch(
          download_file(url_xls, paste0(data_dir,"mmc_mltc/data"), paste0("en", month, "_", year, ".xls")),
          error = function(e) {
            message("Failed for both .xlsx and .xls for ", month, "/", year)
          }
        )
      }
    }
  }
  
  mmc_county_list <- c("ALBANY","ALLEGANY","BROOME","CATTARAUGUS","CAYUGA","CHAUTAUQUA","CHEMUMG","CHEMUNG","CHENANGO",
                       "CLINTON","COLUMBIA","CORTLAND","DELAWARE","DUTCHESS","ERIE","ESSEX","FRANKLIN","FULTON",
                       "GENESEE","GREENE","HAMILTON","HERKIMER","JEFFERSON","LEWIS","LIVINGSTON","MADISON","MONROE",
                       "MONTGOMERY","NASSAU","NEW YORK","NEW YORK CITY","NIAGARA","ONEIDA","ONONDAGA","ONTARIO","ORANGE",
                       "ORLEANS","OSWEGO","OTSEGO","PUTNAM","RENSSELAER","ROCKLAND","SAINT LAWRENCE","SARATOGA","SCHENECTADY", 
                       "SCHOHARIE","SCHUYLER","SENECA","ST LAWRENCE","STEUBEN","SUFFOLK","SULLIVAN","TIOGA","TOMPKINS",
                       "ULSTER","WARREN","WASHINGTON","WAYNE","WESTCHESTER","WYOMING","YATES")
  
  downloaded_mmc_files <- list.files(paste0(data_dir,"mmc_mltc/data"),
                                     recursive = TRUE,
                                     full.names = TRUE,
                                     pattern = "xls")
  
  report_sheets <- downloaded_mmc_files %>% 
    str_replace("en10_24","en10_2024") %>% 
    tibble() %>% 
    rename("report" = 1) %>% 
    filter(str_detect(report,"Medicaid-Program-Enrollment-by-Month-Beginning-2009.csv",negate = TRUE),
           str_detect(report,"Medicaid-Spending-by-Fiscal_Year.csv",negate = TRUE)) %>% 
    rowwise() %>% 
    mutate(sheets = readxl::excel_sheets(report) %>% paste(collapse = ";")) %>% 
    ungroup() %>% 
    separate_rows(sheets,sep = ";") %>% 
    mutate(report_my = report %>%
             str_remove(paste0(data_dir,"mmc_mltc/data/")) %>% 
             str_remove(".xlsx") %>% 
             str_remove(".xls") %>% 
             str_remove("en") %>%
             paste0(.,"_01") %>% 
             as.Date("%m_%Y_%d"))
  
  parse_mltc <- function(input_file){
    
    mltc_tab_name <- readxl::excel_sheets(input_file)
    mltc_tab_name <- mltc_tab_name[str_detect(mltc_tab_name,"Managed Long Term")]
    
    check_rows <- TRUE
    check_rows_i <- 0
    
    while(check_rows){
      if(check_rows_i > 10){
        check_rows <- FALSE
      } else if((readxl::read_excel(input_file,
                                    sheet = mltc_tab_name,
                                    skip = check_rows_i,
                                    n_max = check_rows_i + 1) %>% 
                 names() %>%
                 str_to_upper() %>%
                 str_trim() == "COUNTY") %>%
                sum(na.rm = TRUE) > 0){
        check_rows <- FALSE
      } else{
        check_rows <- TRUE
        check_rows_i <- check_rows_i + 1
      }
    }
    
    check_rows_i <- ifelse(check_rows_i >= 10,0,check_rows_i)
    
    temp <- readxl::read_excel(input_file,
                               sheet = mltc_tab_name,
                               skip = check_rows_i) %>%
      janitor::clean_names()
    
    temp %>% 
      select(1,ncol(temp)) %>% 
      rename("plan_type" = 1,
             "number_of_recipients" = 2) %>% 
      filter(!is.na(number_of_recipients),
             plan_type %>% str_to_lower() %>% str_detect("pace|partial"),
             plan_type %>% str_to_lower() %>% str_detect(" cny",negate = TRUE),
             plan_type %>% str_to_lower() %>% str_detect("hudson",negate = TRUE)) %>%
      mutate(plan_type = case_when(plan_type %>% str_to_lower() %>% str_detect("pace") ~ "PACE",
                                   plan_type %>% str_to_lower() %>% str_detect("partial") ~ "Partial"),
             number_of_recipients = number_of_recipients %>% 
               as.character() %>% 
               str_remove_all(",") %>%
               as.numeric()) %>% 
      group_by(plan_type) %>% 
      summarise(number_of_recipients = max(number_of_recipients)) %>% 
      ungroup() %>% 
      return()
  }
  
  start_time_mltc <- Sys.time()
  mltc <- report_sheets %>%
    filter(sheets %>% str_to_lower() %>% str_detect("long term"),
           report_my >= as.Date("2009-02-01")) %>%
    future_map2_dfr(.x = dplyr::pull(.,report),
                    .y = dplyr::pull(.,report_my),
                    .f = ~ parse_mltc(.x) %>%
                      mutate(month_year = .y),
                    .progress = TRUE)
  end_time_mltc <- Sys.time()
  # end_time_mltc - start_time_mltc
  
  parse_map <- function(input_file){
    
    map_tab_name <- readxl::excel_sheets(input_file)
    map_tab_name <- map_tab_name[str_detect(map_tab_name,"Medicaid Advantage Plus")]
    
    check_rows <- TRUE
    check_rows_i <- 0
    
    while(check_rows){
      if(check_rows_i > 10){
        check_rows <- FALSE
      } else if((readxl::read_excel(input_file,
                                    sheet = map_tab_name,
                                    skip = check_rows_i,
                                    n_max = check_rows_i + 1) %>% 
                 names() %>%
                 str_to_upper() %>%
                 str_trim() == "COUNTY") %>%
                sum(na.rm = TRUE) > 0){
        check_rows <- FALSE
      } else{
        check_rows <- TRUE
        check_rows_i <- check_rows_i + 1
      }
    }
    
    check_rows_i <- ifelse(check_rows_i >= 10,0,check_rows_i)
    
    temp <- readxl::read_excel(input_file,
                               sheet = map_tab_name,
                               skip = check_rows_i) %>%
      janitor::clean_names()
    
    if(ncol(temp) == 3){
      names(temp) <- c("county",
                       "plan_name",
                       "total_enrolled")
      temp <- temp %>% 
        mutate(total_enrolled = total_enrolled %>% 
                 as.character() %>% 
                 str_remove_all(",") %>%
                 as.numeric())
    } else if(ncol(temp) == 2){
      names(temp) <- c("plan_name",
                       "total_enrolled")
      temp <- temp %>%
        mutate(total_enrolled = total_enrolled %>% 
                 as.character() %>% 
                 str_remove_all(",") %>%
                 as.numeric(),
               county = NA)
    } else if(names(temp)[1] == "medicaid_advantage_plus_enrollment_by_county_and_plan"){
      temp <- temp %>% 
        select(1:2) %>% 
        rename("plan_name" = 1,
               "total_enrolled" = 2) %>% 
        mutate(total_enrolled = total_enrolled %>% 
                 as.character() %>% 
                 str_remove_all(",") %>%
                 as.numeric(),
               county = NA)
    }
    
    temp %>%
      filter(str_trim(str_to_lower(county)) %in% c("grand total","state total") |
               str_trim(str_to_lower(plan_name)) %in% c("grand total","state total")) %>%
      rename("number_of_recipients" = "total_enrolled") %>% 
      mutate(plan_type = "MAP") %>%
      select(plan_type,
             number_of_recipients) %>% 
      return()
  }
  
  start_time_map <- Sys.time()
  map <- report_sheets %>%
    filter(sheets %>% str_to_lower() %>% str_detect("advantage plus"),
           report_my >= as.Date("2009-02-01")) %>%
    future_map2_dfr(.x = dplyr::pull(.,report),
                    .y = dplyr::pull(.,report_my),
                    .f = ~ parse_map(.x) %>%
                      mutate(month_year = .y),
                    .progress = TRUE)
  end_time_map <- Sys.time()
  # end_time_map - start_time_map
  
  nys_medicaid <- mltc %>% 
    mutate(eligibility_year = format(month_year,"%Y") %>% as.numeric,
           eligibility_month = format(month_year,"%m") %>% as.numeric) %>% 
    select(eligibility_year,
           eligibility_month,
           plan_type,
           number_of_recipients) %>% 
    bind_rows(map %>% 
                mutate(eligibility_year = format(month_year,"%Y") %>% as.numeric,
                       eligibility_month = format(month_year,"%m") %>% as.numeric) %>% 
                select(eligibility_year,
                       eligibility_month,
                       plan_type,
                       number_of_recipients))
  
  write_csv(nys_medicaid,paste0(data_dir,'mmc_mltc/data/nys_mltc_enrollment.csv'))
} else{
  nys_medicaid <- read_csv(paste0(data_dir,'mmc_mltc/data/nys_mltc_enrollment.csv'))
}
```

```{r setup_medicaid_spending_data_ingestion, include=FALSE}
if(params$use_local_data == FALSE){
  # https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/
  try_downloading_report <- function(){
    # Use GET with a User-Agent header
    response <- GET(pdf_url, user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/121.0.0.0 Safari/537.36"))
    
    # Check response status
    if (status_code(response) == 200) {
      writeBin(content(response, "raw"), pdf_file)
      message("File downloaded successfully.")
    } else {
      message("Failed to download file. HTTP status: ", status_code(response))
    }
  }
  
  # Download Global Cap Quarterly Reports
  for(i in c(2011:(as.numeric(format(Sys.Date(),"%Y"))))){
    for(j in c(1:4)){
      qrt_number <- case_when(j == 1 ~ "1st",
                              j == 2 ~ "2nd",
                              j == 3 ~ "3rd",
                              j == 4 ~ "4th")
      
      pdf_url <- paste0("https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/monthly/sfy_",i,"-",i+1,"/docs/",qrt_number,"_qtr_rpt.pdf")
      
      pdf_file <- paste0(data_dir,"global_cap_reports/nys_medicaid_",i,"-",i+1,"_q",j,"report.pdf")
      
      if(file.exists(pdf_file) == FALSE){
        try_downloading_report()
      } else{
        message("File already downloaded")
      }
    }
  }
  
  # Download Global Cap Monthly Reports
  for(i in c(2011:(as.numeric(format(Sys.Date(),"%Y"))))){
    for(j in c(month.name,month.abb,"Sept")){
      for(k in c(0,1)){
        pdf_url <- paste0("https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/monthly/sfy_",i,"-",i+1,"/docs/",j,"_",i+k,"_report.pdf") %>% 
          str_to_lower()
        
        pdf_file <- paste0(data_dir,"global_cap_reports/nys_medicaid_",i,"-",i+1,"_month_",j,"_",i+k,"_report.pdf")
        
        if(file.exists(pdf_file) == FALSE){
          try_downloading_report()
        } else{
          message("File already downloaded")
        }
      }
    }
  }
  
  report_names <- list.files(paste0(data_dir,"global_cap_reports/"),
                             pattern = ".pdf",
                             full.names = FALSE,
                             recursive = FALSE) %>% 
    # Unusually formatted non-Q4 report
    setdiff("nys_medicaid_2024-2025_q2report.pdf")
  
  report_paths <- list.files(paste0(data_dir,"global_cap_reports/"),
                             pattern = ".pdf",
                             full.names = TRUE,
                             recursive = FALSE) %>% 
    # Unusually formatted non-Q4 report
    setdiff("/Users/isaac/Dropbox/S2P Datasets Shared Folder/MLTC_MMC_NH_data/global_cap_reports//nys_medicaid_2024-2025_q2report.pdf")
  
  for(i in c(1:length(report_names))){
    report_name <- report_names[i]
    
    # Extract text from the PDF
    pdf_text <- pdf_text(report_paths[i])
    
    report_table_page <- c(1:length(pdf_text))[pdf_text %>% str_to_lower() %>% str_detect("actual") &
                                                 pdf_text %>% str_to_lower() %>% str_detect("variance") &
                                                 (pdf_text %>% str_to_lower() %>% str_detect("category of service") | 
                                                    pdf_text %>% str_to_lower() %>% str_detect("category of spending") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 1") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 2") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 3") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 4")) &
                                                 pdf_text %>% str_detect("TOTAL") &
                                                 pdf_text %>% str_to_lower() %>% str_detect("table of contents",negate = TRUE)] %>% 
      min()
    
    pdf_table <- pdf_text[report_table_page]
    
    dollar_multiplier <- case_when(pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("in millions") ~ 1000000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("in thousands") ~ 1000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("millions") ~ 1000000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("thousands") ~ 1000,
                                   TRUE ~ 1000000)
    
    # Parse the table from the extracted text
    table_data <- str_split(pdf_table,"\n")[[1]] %>% 
      # Clean and structure the data
      str_trim() %>%
      as_tibble()
    
    table_start_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_to_lower() %>% str_sub(1,19) == "category of service" |
               value %>% str_to_lower() %>% str_sub(1,20) == "category of spending" |
               (value %>% str_to_lower() %>% str_detect("estimated") &
                  value %>% str_to_lower() %>% str_detect("actual") &
                  value %>% str_to_lower() %>% str_detect("variance"))) %>% 
      dplyr::pull(rowid)
    
    first_data_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_count("\\$") == 3,
             rowid > table_start_row) %>% 
      dplyr::pull(rowid) %>%
      min()
    
    table_end_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_sub(1,5) == "TOTAL") %>% 
      dplyr::pull(rowid)
    
    header_rows <- table_data %>% 
      slice(table_start_row:first_data_row - 1) %>% 
      as.character() %>% 
      paste(collapse = " ") %>% 
      str_to_lower()
    
    table_col_names <- case_when(str_detect(header_rows,"target") ~ c("category_of_service","target","actual","variance"),
                                 str_detect(header_rows,"estimated") ~ c("category_of_service","estimated","actual","variance"))
    
    table_clean <- table_data %>%
      slice(first_data_row:table_end_row) %>% 
      separate(col = value, into = c("Column1", "Column2", "Column3", "Column4"), sep = "\\s{2,}", fill = "right")
    
    names(table_clean) <- table_col_names
    
    table_clean <- table_clean %>% 
      mutate(report = report_name %>% str_remove(".pdf"),
             multiplier = dollar_multiplier)
    
    if(i == 1){
      nys_medicaid_spending <- table_clean
    } else{
      nys_medicaid_spending <- bind_rows(nys_medicaid_spending,table_clean)
    }
    paste0(i," of ",length(report_names),"(",scales::percent(i/length(report_names),accuracy = 0.1),") | ",report_name) %>% print()
  }
  write_csv(nys_medicaid_spending,paste0(data_dir,'mmc_mltc/data/Medicaid-Spending-by-Fiscal_Year.csv'))
} else{
  nys_medicaid_spending <- read_csv(paste0(data_dir,'mmc_mltc/data/Medicaid-Spending-by-Fiscal_Year.csv'))
}
```

```{r setup_medicaid_enrollment_vs_spending_analysis, include=FALSE}
mltc_analysis <- nys_medicaid %>% 
  # filter(managed_care_vs_fee_for_service == "MMC",
  #        plan_type %in% c("PACE",
  #                         "PARTIAL MLTC",
  #                         "MAP")) %>% 
  mutate(month_year = paste(eligibility_year,eligibility_month,"01",sep = "-") %>% as.Date(),
         year = paste(eligibility_year,"01","01",sep = "-") %>% as.Date())

mmc_global_cap_analysis <- nys_medicaid_spending %>% 
  mutate(reference = ifelse(is.na(target),estimated,target),
         ref_type = ifelse(is.na(target),"estimated","target"),
         across(.cols = c("reference","actual"),
                .fns = ~ (.x %>%
                            # Remove $ and ,
                            str_remove_all("[\\$,]") %>%
                            # Replace parentheses with a negative sign
                            str_replace_all("\\((.*)\\)", "-\\1") %>%
                            # Convert to numeric
                            as.numeric()) * multiplier),
         difference = actual - reference) %>% 
  select(category_of_service,
         ref_type,
         reference,
         actual,
         report) %>% 
  # Only include year-end reports in the analysis
  filter(str_detect(report,"q4|Mar")) %>% 
  mutate(fiscal_yr = str_sub(report,14,22),
         type = case_when(category_of_service %in% c("Long Term Managed Care") ~ "Long Term Care")) %>% 
  select(-report)

mmc_global_cap_analysis_data_v2 <- mmc_global_cap_analysis %>% 
  mutate(fiscal_yr_dt = fiscal_yr %>% str_sub(1,4) %>% paste("01","01",sep="-") %>% as.Date()) %>% 
  select(-ref_type,
         -reference) %>% 
  filter(!is.na(type)) %>% 
  group_by(fiscal_yr,
           fiscal_yr_dt,
           type) %>%
  summarise(actual = sum(actual)) %>% 
  ungroup()

mltc_enrollment_vs_spending_table_data <- mltc_analysis %>% 
  mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                               TRUE ~ eligibility_year)) %>% 
  group_by(plan_type,
           fiscal_yr) %>% 
  summarise(enrollee_months = sum(number_of_recipients)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = plan_type,
              values_from = enrollee_months) %>% 
  left_join(mltc_analysis %>% 
              mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                                           TRUE ~ eligibility_year)) %>% 
              group_by(fiscal_yr,eligibility_month) %>% 
              summarise(enrollee_months = sum(number_of_recipients)) %>% 
              ungroup() %>% 
              group_by(fiscal_yr) %>% 
              slice_max(eligibility_month) %>% 
              ungroup() %>% 
              select(fiscal_yr,enrollee_months) %>% 
              rename("end_of_period_enrollment" = "enrollee_months"),
            by = "fiscal_yr") %>% 
  left_join(mmc_global_cap_analysis_data_v2 %>% 
              filter(type == "Long Term Care") %>% 
              select(fiscal_yr,
                     actual) %>% 
              mutate(fiscal_yr = fiscal_yr %>% str_sub(1,4) %>% as.numeric()) %>% 
              rename("mltc_spending" = "actual"),
            by = "fiscal_yr") %>% 
  left_join(mltc_analysis %>%
              mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                                           TRUE ~ eligibility_year)) %>%
              group_by(fiscal_yr,eligibility_month) %>%
              summarise(enrollee_months = sum(number_of_recipients)) %>%
              ungroup() %>%
              group_by(fiscal_yr) %>%
              summarise(n_months = n(),
                        mean_number_of_enrollees = mean(enrollee_months)) %>%
              ungroup(),
            by = "fiscal_yr") %>% 
  left_join(mltc_analysis %>% 
              mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                                           TRUE ~ eligibility_year)) %>% 
              group_by(fiscal_yr,eligibility_month) %>% 
              summarise(enrollee_months = sum(number_of_recipients)) %>% 
              ungroup() %>% 
              arrange(desc(fiscal_yr),
                      desc(eligibility_month)) %>% 
              mutate(enrollee_months_delta = enrollee_months  - lead(enrollee_months),
                     enrollee_months_delta_pct = enrollee_months_delta / lead(enrollee_months,n=1)) %>% 
              group_by(fiscal_yr) %>% 
              summarise(mean_monthly_change_pct = mean(enrollee_months_delta_pct)) %>% 
              ungroup(),
            by = "fiscal_yr") %>% 
  arrange(desc(fiscal_yr)) %>% 
  filter(fiscal_yr >= 2014) %>% 
  rename("PARTIAL MLTC" = "Partial") %>% 
  mutate(fiscal_yr_dt = fiscal_yr %>% str_sub(1,4) %>% paste("01","01",sep="-") %>% as.Date(),
         fiscal_yr = fiscal_yr_dt %>% format("%Y") %>% as.numeric(),
         fiscal_yr = paste(fiscal_yr,str_sub(fiscal_yr + 1,3,4),sep = "-"),
         fiscal_yr = paste0("FY",str_sub(fiscal_yr,-2)),
         enrollee_months = MAP + PACE + `PARTIAL MLTC`,
         mltc_spending_delta = mltc_spending  - lead(mltc_spending),
         mltc_spending_delta_pct = mltc_spending_delta / lead(mltc_spending,n=1),
         spending_per_enrollee_month = mltc_spending / enrollee_months,
         spending_per_enrollee_month_delta = spending_per_enrollee_month  - lead(spending_per_enrollee_month),
         spending_per_enrollee_month_delta_pct = spending_per_enrollee_month_delta / lead(spending_per_enrollee_month,n=1),
         enrollee_month_delta = enrollee_months - lead(enrollee_months),
         enrollee_month_delta_pct = enrollee_month_delta / lead(enrollee_months,n=1),
         mean_spend_per_enrollee = mltc_spending / mean_number_of_enrollees,
         mean_spend_per_enrollee_delta = mean_spend_per_enrollee - lead(mean_spend_per_enrollee),
         mean_spend_per_enrollee_delta_pct = mean_spend_per_enrollee_delta / lead(mean_spend_per_enrollee,n=1),
         across(.cols = c(enrollee_month_delta,enrollee_month_delta_pct),
                .fn = ~ ifelse(n_months != 12,NA,.x))) %>% 
  select(fiscal_yr,
         n_months,
         MAP,
         PACE,
         `PARTIAL MLTC`,
         enrollee_months,
         end_of_period_enrollment,
         mean_monthly_change_pct,
         enrollee_month_delta,
         enrollee_month_delta_pct,
         mltc_spending,
         mltc_spending_delta,
         mltc_spending_delta_pct,
         mean_number_of_enrollees,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta,
         mean_spend_per_enrollee_delta_pct)

if(mltc_enrollment_vs_spending_table_data %>% slice(1) %>% dplyr::pull(n_months) < 12){
  projection_data_ytd <- mltc_enrollment_vs_spending_table_data %>% slice(1)
  projection_data_base_year <- mltc_enrollment_vs_spending_table_data %>% slice(2)
  projection_data_base_5_years <- mltc_enrollment_vs_spending_table_data %>% slice(2:6)
  
  fiscal_yr_projected <- paste0(projection_data_ytd$fiscal_yr," (Estimate)")
  
  enrollee_months_projected <- projection_data_ytd$enrollee_months / (projection_data_ytd$n_months / 12)
  map_projected <- projection_data_ytd$MAP / (projection_data_ytd$n_months / 12)
  pace_projected <- projection_data_ytd$PACE / (projection_data_ytd$n_months / 12)
  partial_projected <- projection_data_ytd$`PARTIAL MLTC` / (projection_data_ytd$n_months / 12)
  end_of_period_enrollment_projected <- enrollee_months_projected / 12
  mean_number_of_enrollees_projected <- ((enrollee_months_projected + projection_data_base_year$enrollee_months) / 2) / 12
  enrollee_month_delta_projected <- enrollee_months_projected - projection_data_base_year$enrollee_months
  enrollee_month_delta_pct_projected <- enrollee_month_delta_projected / projection_data_base_year$enrollee_months
  
  spending_growth_rate_projected <- mean(projection_data_base_5_years$mean_spend_per_enrollee_delta_pct)
  mean_spend_per_enrollee_projected <- (1 + spending_growth_rate_projected) * projection_data_base_year$mean_spend_per_enrollee
  mltc_spending_projected <- mean_number_of_enrollees_projected * mean_spend_per_enrollee_projected
  mltc_spending_delta_projected <- mltc_spending_projected - projection_data_base_year$mltc_spending
  mltc_spending_delta_pct_projected <- mltc_spending_delta_projected / projection_data_base_year$mltc_spending
  mean_spend_per_enrollee_delta_projected <- mean_spend_per_enrollee_projected - projection_data_base_year$mean_spend_per_enrollee
  mean_spend_per_enrollee_delta_pct_projected <- mean_spend_per_enrollee_delta_projected / projection_data_base_year$mean_spend_per_enrollee
  
  mltc_enrollment_vs_spending_projection_table_data <- tibble(
    fiscal_yr = fiscal_yr_projected,
    MAP = map_projected,
    PACE = pace_projected,
    `PARTIAL MLTC` = partial_projected,
    enrollee_months = enrollee_months_projected,
    enrollee_month_delta = enrollee_month_delta_projected,
    enrollee_month_delta_pct = enrollee_month_delta_pct_projected,
    end_of_period_enrollment = end_of_period_enrollment_projected,
    mltc_spending = mltc_spending_projected,
    mltc_spending_delta = mltc_spending_delta_projected,
    mltc_spending_delta_pct = mltc_spending_delta_pct_projected,
    mean_number_of_enrollees = mean_number_of_enrollees_projected,
    mean_spend_per_enrollee = mean_spend_per_enrollee_projected,
    mean_spend_per_enrollee_delta = mean_spend_per_enrollee_delta_projected,
    mean_spend_per_enrollee_delta_pct = mean_spend_per_enrollee_delta_pct_projected
  )
} else{
  mltc_enrollment_vs_spending_projection_table_data <- tibble()
}

mltc_enrollment_vs_spending_table <- bind_rows(mltc_enrollment_vs_spending_projection_table_data,
                                               mltc_enrollment_vs_spending_table_data %>% 
                                                 mutate(fiscal_yr = ifelse(n_months < 12,paste0(fiscal_yr," (YTD)"),fiscal_yr))) %>% 
  relocate(any_of(names(mltc_enrollment_vs_spending_table_data))) %>% 
  gt() %>% 
  tab_header(title = md("**Managed Long Term Care Enrollment and Spending, by Fiscal Year**"),
             subtitle = paste0("New York State (",
                               # mmc_global_cap_analysis_data_v2 %>% filter(type == "Long Term Care") %>% slice_min(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                               # " to ",
                               # mmc_global_cap_analysis_data_v2 %>% filter(type == "Long Term Care") %>% slice_max(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                               "Since April 2014",
                               ")")) %>%
  fmt_integer(c(n_months,
                MAP,
                PACE,
                `PARTIAL MLTC`,
                enrollee_months,
                enrollee_month_delta,
                mean_number_of_enrollees,
                end_of_period_enrollment)) %>% 
  fmt_currency(c(mltc_spending,
                 mltc_spending_delta,
                 mean_spend_per_enrollee,
                 mean_spend_per_enrollee_delta),
               decimals = 0) %>% 
  fmt_percent(c(enrollee_month_delta_pct,
                mltc_spending_delta_pct,
                mean_spend_per_enrollee_delta_pct,
                mean_monthly_change_pct),
              decimals = 1) %>% 
  tab_spanner(label = "Enrollee Months, by Plan Type",
              columns = c(MAP,
                          PACE,
                          `PARTIAL MLTC`)) %>% 
  tab_spanner(label = "Total Enrollment",
              columns = c(enrollee_months,
                          mean_monthly_change_pct,
                          enrollee_month_delta,
                          enrollee_month_delta_pct,
                          end_of_period_enrollment)) %>% 
  tab_spanner(label = "State Spending",
              columns = c(mltc_spending,
                          mltc_spending_delta,
                          mltc_spending_delta_pct)) %>% 
  tab_spanner(label = "Average Annual Spending per Enrollee",
              columns = c(mean_number_of_enrollees,
                          mean_spend_per_enrollee,
                          mean_spend_per_enrollee_delta,
                          mean_spend_per_enrollee_delta_pct)) %>% 
  cols_label(fiscal_yr = "Fiscal Year",
             n_months = "Months Reported",
             `PARTIAL MLTC` = "MLTC",
             enrollee_months = "Enrollee Months",
             mean_monthly_change_pct = "Average Monthly Change %",
             enrollee_month_delta = "Annual Change",
             enrollee_month_delta_pct = "Annual Change %",
             end_of_period_enrollment = "End of Period Enrollment",
             mltc_spending = "Spending",
             mltc_spending_delta = "Annual Change",
             mltc_spending_delta_pct = "Annual Change %",
             mean_number_of_enrollees = "Average Monthly Number of Enrollees",
             mean_spend_per_enrollee = "Spending",
             mean_spend_per_enrollee_delta = "Annual Change",
             mean_spend_per_enrollee_delta_pct = "Annual Change %") %>% 
  cols_align(align = "left", columns = fiscal_yr) %>% 
  tab_style(style = cell_text(weight = "bold"),
            locations = list(cells_column_labels(),
                             cells_column_spanners())) %>% 
  tab_source_note(source_note = md("**Data Sources:** New York State Department of Health Medicaid Managed Care Enrollment Reports, and Medicaid Global Spending Cap Updates")) %>% 
  tab_source_note(source_note = md(paste0("**Data as of:** ",format(Sys.Date(),"%B %d, %Y")))) %>% 
  tab_options(table.font.size = px(10),
              table.font.names = "Helvetica",
              data_row.padding = px(2),
              table.width = pct(95),
              container.width = px(1000))

# Calculate scaling factors for normalization
max_enrollee_mltc <- max(mltc_enrollment_vs_spending_table_data$enrollee_months,na.rm=TRUE) / 1e6
max_spending_mltc <- max(mltc_enrollment_vs_spending_table_data$mltc_spending,na.rm=TRUE) / 1e9

mltc_enrollment_vs_spending_graph <- mltc_enrollment_vs_spending_table_data %>% 
  select(fiscal_yr,
         enrollee_months,
         mltc_spending) %>% 
  # Create the plot
  ggplot(aes(x = fiscal_yr)) +
  # Plot enrollee_months as bars
  geom_bar(aes(y = enrollee_months / 1e6, fill = "Enrollment"),
           stat = "identity", alpha = 0.9) +
  # Plot mltc_spending as a line with points
  geom_line(aes(y = (mltc_spending / 1e9) * (max_enrollee_mltc / max_spending_mltc),
                group = 1, color = "Spending")) +
  geom_point(aes(y = (mltc_spending / 1e9) * (max_enrollee_mltc / max_spending_mltc)), color = "black") +
  # Left y-axis for enrollee_months
  scale_y_continuous(
    name = "Enrollee Months (millions)\n",
    labels = scales::label_comma(),
    breaks = breaks_pretty(n = 8),
    sec.axis = sec_axis(~ . * (max_spending_mltc / max_enrollee_mltc), 
                        name = "MLTC Spending\n", 
                        breaks = breaks_pretty(n = 8),
                        labels = scales::label_dollar(scale = 1, suffix = "B"))) +
  scale_fill_manual(values = "#B9CDE5") +
  scale_color_manual(values = "black") +
  # Formatting
  labs(x = "\nFiscal Year",
       title = "Medicaid Managed Long Term Care Enrollment and Spending, by Fiscal Year",
       subtitle = paste0("New York State (",
                         # mmc_global_cap_analysis_data_v2 %>% filter(type == "Long Term Care") %>% slice_min(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                         # " to ",
                         # mmc_global_cap_analysis_data_v2 %>% filter(type == "Long Term Care") %>% slice_max(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                               "Since April 2014",
                         ")"),
       color = "",
       fill = "",
       caption = paste0("Data Sources: New York State Department of Health Medicaid Managed Care Enrollment Reports, and Medicaid Global Spending Cap Updates\nData as of: ",format(Sys.Date(),"%B %d, %Y"))) +
  theme_light(base_size = 16) +
  isaac_step2_graph_theme +
  theme(legend.text = element_text(face = "bold"))

# mltc_trends_by_plan_name <- mltc_analysis %>% 
#   group_by(plan_name,month_year,plan_type) %>% 
#   summarise(enrolled = sum(number_of_recipients)) %>% 
#   ungroup() %>% 
#   # mutate(plan_name = plan_name %>% factor(levels = mltc_analysis %>% 
#   #                                           group_by(plan_name) %>% 
#   #                                           summarise(enrolled = sum(number_of_recipients)) %>% 
#   #                                           ungroup() %>% 
#   #                                           arrange(desc(enrolled)) %>% 
#   #                                           dplyr::pull(plan_name))) %>% 
#   ggplot(aes(x = month_year,
#              y = enrolled,
#              color = plan_type)) + 
#   geom_line(linewidth = 1.25) +
#   labs(title = "Medicaid Managed Long Term Care Plans, Statewide Enrollment by Plan Name, Plan Type, and Month",
#        subtitle = paste0("New York State (",
#                          format(min(mltc_analysis$month_year),"%b %Y"),
#                          " to ",
#                          format(max(mltc_analysis$month_year),"%b %Y"),
#                          ")"),
#        x = "",
#        y = "Number of Enrollees\n",
#        color = "Plan Type",
#        caption = paste0("Data Source: Health Data NY\nData as of: ",format(Sys.Date(),"%B %d, %Y"))) +
#   scale_x_date(breaks = breaks_pretty(n = 4)) +
#   scale_y_continuous(labels = comma_format(accuracy = 1),
#                      limits = c(0,NA)) +
#   scale_color_manual(values = c("#1b9e77","#d95f02","#7570b3")) +
#   theme_light(base_size = 12) +
#   isaac_step2_graph_theme +
#   facet_wrap(~plan_name,
#              labeller = label_wrap_gen(width = 20),
#              ncol = 5,
#              axes = "all",
#              scales = "free_y")

# mltc_trends_by_plan_name_table <- mltc_analysis %>% 
#   group_by(plan_name,year,plan_type) %>% 
#   summarise(enrolled = sum(number_of_recipients)) %>% 
#   ungroup() %>% 
#   mutate(year = year %>% format("%Y") %>% as.numeric()) %>% 
#   pivot_wider(names_from = year,
#               values_from = enrolled,
#               names_sort = TRUE,
#               values_fill = 0) %>% 
#   relocate(plan_type,plan_name) %>% 
#   arrange(plan_type,plan_name) %>% 
#   gt(groupname_col = "plan_type",
#      row_group_as_column = TRUE) %>% 
#   tab_header(title = md("**Medicaid Managed Long Term Care Plan<br>Member Months, by Plan Type, Plan Name, and Year**"),
#              subtitle = paste0("New York State (",
#                                format(min(mltc_analysis$month_year),"%b %Y"),
#                                " to ",
#                                format(max(mltc_analysis$month_year),"%b %Y"),
#                                ")")) %>% 
#   opt_row_striping() %>% 
#   fmt_integer(is.numeric) %>% 
#   cols_label(plan_name = "Plan Name",
#              plan_type = "Plan Type") %>% 
#   cols_align(align = "left", columns = plan_name) %>% 
#   tab_style(style = cell_text(weight = "bold"),
#             locations = cells_column_labels()) %>% 
#   tab_source_note(source_note = md("Scroll right to see additional years")) %>% 
#   tab_source_note(source_note = md("**Data Source:** Health Data NY")) %>% 
#   tab_source_note(source_note = md(paste0("**Data as of:** ",format(Sys.Date(),"%B %d, %Y")))) %>% 
#   tab_options(table.font.size = px(10),
#               table.font.names = "Helvetica",
#               data_row.padding = px(2))

mltc_enrollment_vs_spending_table_data %>%
  select(fiscal_yr,
         n_months,
         MAP,
         PACE,
         `PARTIAL MLTC`,
         enrollee_months,
         mean_number_of_enrollees,
         end_of_period_enrollment,
         mean_monthly_change_pct,
         enrollee_month_delta,
         enrollee_month_delta_pct,
         mltc_spending,
         mltc_spending_delta,
         mltc_spending_delta_pct,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta,
         mean_spend_per_enrollee_delta_pct) %>% 
  mutate(fiscal_yr = ifelse(n_months != 12,paste0(fiscal_yr," (YTD)"),fiscal_yr),
         across(.cols = -c(fiscal_yr,
                           n_months,
                           MAP,
                           PACE,
                           `PARTIAL MLTC`,
                           enrollee_months,
                           mean_number_of_enrollees,
                           end_of_period_enrollment,
                           mean_monthly_change_pct),
                .fns = ~ ifelse(n_months != 12,NA,.x))) %>% 
  rename("State Fiscal Year" = "fiscal_yr",
         "Months Reported" = "n_months",
         "MAP Enrollee Months" = "MAP",
         "PACE Enrollee Months" = "PACE",
         "PARTIAL MLTC Enrollee Months" = "PARTIAL MLTC",
         "Total Enrollee Months" = "enrollee_months",
         "Average Annual Enrollment" = "mean_number_of_enrollees",
         "End of Period Enrollment" = "end_of_period_enrollment",
         "Enrollee Months Average Monthly % Growth" = "mean_monthly_change_pct",
         "Enrollee Months Annual Change" = "enrollee_month_delta",
         "Enrollee Months % Growth" = "enrollee_month_delta_pct",
         "State Spending (Millions)" = "mltc_spending",
         "State Spending Annual Change (Millions)" = "mltc_spending_delta",
         "State Spending Annual Change %" = "mltc_spending_delta_pct",
         "Average Annual Spending per Enrollee" = "mean_spend_per_enrollee",
         "Average Annual Spending per Enrollee Annual Change" = "mean_spend_per_enrollee_delta",
         "Average Annual Per Capita Spending % Growth" = "mean_spend_per_enrollee_delta_pct") %>% 
  write_csv("mltc_table_data.csv")

mltc_enrollment_vs_spending_table_data %>%
  filter(n_months == 12) %>% 
  mutate(data_type = "Actual",
         fiscal_yr_n = fiscal_yr %>% str_remove("FY") %>% as.numeric(),
         fiscal_yr = ifelse(n_months != 12,paste0(fiscal_yr," (YTD)"),fiscal_yr)) %>% 
  arrange(fiscal_yr) %>% 
  mutate(mean_number_of_enrollees_delta_pct = (mean_number_of_enrollees - lag(mean_number_of_enrollees)) / lag(mean_number_of_enrollees)) %>% 
  select(data_type,
         fiscal_yr,
         mltc_spending,
         mean_number_of_enrollees,
         mean_number_of_enrollees_delta_pct,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta_pct) %>% 
  rename("Data Type" = "data_type",
         "Fiscal Year" = "fiscal_yr",
         "State Spending" = "mltc_spending",
         "Average Annual Enrollment" = "mean_number_of_enrollees",
         "Average Annual Enrollment Change %" = "mean_number_of_enrollees_delta_pct",
         "Average Annual Spending per Enrollee" = "mean_spend_per_enrollee",
         "Average Annual Spending per Enrollee Annual Change %" = "mean_spend_per_enrollee_delta_pct") %>% 
  write_csv("mltc_projection_data.csv")
```

```{r setup_excel_workbook_data, include=FALSE}
datatable_page <- mltc_enrollment_vs_spending_table_data %>% 
  select(fiscal_yr,
         n_months,
         MAP,
         PACE,
         `PARTIAL MLTC`,
         enrollee_months,
         mean_number_of_enrollees,
         end_of_period_enrollment,
         enrollee_month_delta,
         enrollee_month_delta_pct,
         mltc_spending,
         mltc_spending_delta,
         mltc_spending_delta_pct,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta,
         mean_spend_per_enrollee_delta_pct)

projection_data_page <- mltc_enrollment_vs_spending_table_data %>% 
  select(fiscal_yr,
         mltc_spending,
         mean_number_of_enrollees,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta_pct) %>% 
  arrange(desc(fiscal_yr)) %>% 
  mutate(mean_number_of_enrollees_delta = mean_number_of_enrollees - lead(mean_number_of_enrollees),
         mean_number_of_enrollees_delta_pct = mean_number_of_enrollees_delta / lead(mean_number_of_enrollees,n=1)) %>% 
  select(fiscal_yr,
         mltc_spending,
         mean_number_of_enrollees,
         mean_number_of_enrollees_delta_pct,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta_pct) %>% 
  arrange(fiscal_yr)
```

<br>

### Enrollment and Spending

```{r mltc_enrollment_vs_spending_table}
mltc_enrollment_vs_spending_table %>% 
  tab_source_note(
    # mltc_enrollment_vs_spending_table_data %>%
    #   mutate(across(.cols = ends_with("pct"),
    #                 .fns = ~ percent(.x,accuracy = 0.1)),
    #          across(.cols = contains("spend") & !ends_with("pct"),
    #                 .fns = ~ dollar(.x,accuracy = 1)),
    #          across(.cols = c(MAP,PACE,`PARTIAL MLTC`,
    #                           enrollee_months,enrollee_month_delta,mean_number_of_enrollees),
    #                 .fns = ~ comma(.x,accuracy = 1))) %>% 
    #   rename("Fiscal Year" = "fiscal_yr",
    #          "Months Reported" = "n_months",
    #          "Enrollee Months" = "enrollee_months",
    #          "Average Monthly Change %" = "mean_monthly_change_pct",
    #          "Enrollee Months Annual Change" = "enrollee_month_delta",
    #          "Enrollee Months Annual Change %" = "enrollee_month_delta_pct",
    #          "State Spending" = "mltc_spending",
    #          "State Spending Annual Change" = "mltc_spending_delta",
    #          "State Spending Annual Change %" = "mltc_spending_delta_pct",
    #          "Average Monthly Number of Enrollees" = "mean_number_of_enrollees",
    #          "Average Annual Spending per Enrollee" = "mean_spend_per_enrollee",
    #          "Average Annual Spending per Enrollee Annual Change" = "mean_spend_per_enrollee_delta",
    #          "Average Annual Spending per Enrollee Annual Change %" = "mean_spend_per_enrollee_delta_pct") %>% 
    #   download_this(
    #     output_name = paste0("mltc_enrollment_vs_spending__",str_replace_all(Sys.Date(),"-","_")),
    #     output_extension = ".xlsx",
    #     button_label = "Download Excel",
    #     button_type = "default",
    #   )
    download_link(
      link = "https://github.com/steptwopolicy/s2p_data_analyses/raw/refs/heads/main/mltc_enrollment_vs_spending.xlsx",
      button_label = "Download Interactive Excel Workbook",
      button_type = "default",
      class = "button_download"
    )
  )
```

The interactive Excel workbook allows users to enter growth-rate predictions, and model projections.  
**Data Sources:** New York State Department of Health [Medicaid Managed Care Enrollment Reports](https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/), and [Medicaid Global Spending Cap Updates
](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/)

<br>

```{r mltc_enrollment_vs_spending_graph}
mltc_enrollment_vs_spending_graph
```
**Data Sources:** New York State Department of Health [Medicaid Managed Care Enrollment Reports](https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/), and [Medicaid Global Spending Cap Updates
](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/)

<br>

<!-- ### Enrollment, by Plan Name -->

<!-- ```{r mltc_trends_by_plan_name_table} -->
<!-- mltc_trends_by_plan_name_table -->
<!-- ``` -->
<!-- **Data Sources:** [Health Data NY](https://health.data.ny.gov/Health/Medicaid-Program-Enrollment-by-Month-Beginning-200/m4hz-kzn3/about_data); [New York State Department of Health](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/) -->

<!-- <br> -->

<!-- ```{r mltc_trends_by_plan_name,fig.height=20} -->
<!-- mltc_trends_by_plan_name -->
<!-- ``` -->
<!-- **Data Sources:** [Health Data NY](https://health.data.ny.gov/Health/Medicaid-Program-Enrollment-by-Month-Beginning-200/m4hz-kzn3/about_data); [New York State Department of Health](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/) -->

<!-- <br> -->

***

### Technical Notes

1. <i>MLTC</i> = Partial cap managed long term care.
2. Spending data in this analysis are based on the amounts reported under the 'Long Term Managed Care' category in the New York State Medicaid Global Cap Reports. This category reflects state-reported expenditures specific to Medicaid Managed Long-Term Care programs.
3. Data for FY12 through FY14 are not presented in the enrollment vs. spending analysis because no spending was reported on the 'Long Term Managed Care' line in the New York State Medicaid Global Cap Reports during those years. As a result, the analysis begins with FY15, the first year for which spending on that line was reported.
  
### Definitions

* __Fiscal Year:__ April through March  
* __Enrollee Months:__ The total number of Medicaid managed long term care enrollee months reported within the fiscal year.  
* __End of Period Enrollment:__ The total number of Medicaid managed long term care enrollees at the end of the reporting period. For completed fiscal years, this represents the enrollment total on the final day of the year. For partial fiscal years, it reflects the year-to-date (YTD) enrollment total as of the most recent reporting date.  
* __Average Monthly Change %:__ For each row's respective fiscal year, this column calculates each month's enrollment delta percentage (i.e., _[(that month's enrollment) - (previous month's enrollment)] / (previous month's enrollment))_ and then reports the mean of those twelve percentages.  
* __Enrollee Months Annual Change:__ The difference in the total number of enrollee months between the current fiscal year and the prior fiscal year.  
* __Enrollee Months Annual Change %:__ The percentage increase or decrease in the total enrollee months compared to the prior fiscal year.  
* __State Spending:__ The total Medicaid managed long term care expenditures by the State government.  
* __State Spending Annual Change:__ The year-over-year change in State spending on Medicaid managed long term care.  
* __State Spending Annual Change %:__ The percentage change in the State’s share of Medicaid managed long term care spending compared to the prior fiscal year.  
* __Average Annual Spending per Enrollee:__ The average amount of Medicaid managed long term care State spending per enrollee for the fiscal year.  
* __Average Annual Spending per Enrollee Annual Change:__ The year-over-year change in the average amount of State spending per enrollee, expressed in dollars.  
* __Average Annual Spending per Enrollee Annual Change %:__ The percentage growth in State Medicaid spending per-capita compared to the prior fiscal year.  
