---
output:
  html_document:
    toc: false
params:
  project_directory: "/Users/isaac/Dropbox/S2P Datasets Shared Folder/MLTC_MMC_NH_data/"
  use_local_data: FALSE
---

<style>
.button_download {
font-size: 16px;
font-family: "Arial";
color: black;
background-color: #C7C9D8;
}
</style>

```{r setup_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, error = TRUE, fig.width = 12, fig.height = 8, cache = FALSE)
```

```{r setup_libraries, include=FALSE}
library(tidyverse)
library(scales)
library(gt)
library(pdftools)
library(httr)
library(downloadthis)
data_dir <- params$project_directory
```

```{r isaac_step2_graph_theme, include=FALSE}
isaac_step2_graph_theme <- theme(plot.title.position = "plot",
                                 plot.title = element_text(face = "bold"),
                                 legend.position = "top",
                                 strip.text = element_text(color = "black"),
                                 legend.title = element_text(face = "bold",hjust = 0.5),
                                 axis.title.y = element_text(face = "bold",angle = 90),
                                 axis.title.x = element_text(face = "bold"))
```

```{r setup_nh_data_ingestion, include=FALSE}
if(params$use_local_data == FALSE){
  # https://health.data.ny.gov/Health/Nursing-Home-Weekly-Bed-Census-Beginning-2009/uhyy-xp9s/about_data
  nh_historic <- read_csv("https://health.data.ny.gov/api/views/uhyy-xp9s/rows.csv?accessType=DOWNLOAD") %>% janitor::clean_names()
  write_csv(nh_historic,paste0(data_dir,'nh/data/NH_Weekly_Bed_Census_Beginning_2009.csv'))
} else{
  nh_historic <- read_csv(paste0(data_dir,'nh/data/NH_Weekly_Bed_Census_Beginning_2009.csv'))
}
```

```{r setup_medicaid_spending_data_ingestion, include=FALSE}
if(params$use_local_data == FALSE){
  # https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/
  try_downloading_report <- function(){
    # Use GET with a User-Agent header
    response <- GET(pdf_url, user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/121.0.0.0 Safari/537.36"))
    
    # Check response status
    if (status_code(response) == 200) {
      writeBin(content(response, "raw"), pdf_file)
      message("File downloaded successfully.")
    } else {
      message("Failed to download file. HTTP status: ", status_code(response))
    }
  }
  
  # Download Global Cap Quarterly Reports
  for(i in c(2011:(as.numeric(format(Sys.Date(),"%Y"))))){
    for(j in c(1:4)){
      qrt_number <- case_when(j == 1 ~ "1st",
                              j == 2 ~ "2nd",
                              j == 3 ~ "3rd",
                              j == 4 ~ "4th")
      
      pdf_url <- paste0("https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/monthly/sfy_",i,"-",i+1,"/docs/",qrt_number,"_qtr_rpt.pdf")
      
      pdf_file <- paste0(data_dir,"global_cap_reports/nys_medicaid_",i,"-",i+1,"_q",j,"report.pdf")
      
      if(file.exists(pdf_file) == FALSE){
        try_downloading_report()
      } else{
        message("File already downloaded")
      }
    }
  }
  
  # Download Global Cap Monthly Reports
  for(i in c(2011:(as.numeric(format(Sys.Date(),"%Y"))))){
    for(j in c(month.name,month.abb,"Sept")){
      for(k in c(0,1)){
        pdf_url <- paste0("https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/monthly/sfy_",i,"-",i+1,"/docs/",j,"_",i+k,"_report.pdf") %>% 
          str_to_lower()
        
        pdf_file <- paste0(data_dir,"global_cap_reports/nys_medicaid_",i,"-",i+1,"_month_",j,"_",i+k,"_report.pdf")
        
        if(file.exists(pdf_file) == FALSE){
          try_downloading_report()
        } else{
          message("File already downloaded")
        }
      }
    }
  }
  
  report_names <- list.files(paste0(data_dir,"global_cap_reports/"),
                             pattern = ".pdf",
                             full.names = FALSE,
                             recursive = FALSE) %>% 
    # Unusually formatted non-Q4 report
    setdiff("nys_medicaid_2024-2025_q2report.pdf")
  
  report_paths <- list.files(paste0(data_dir,"global_cap_reports/"),
                             pattern = ".pdf",
                             full.names = TRUE,
                             recursive = FALSE) %>% 
    # Unusually formatted non-Q4 report
    setdiff("/Users/isaac/Dropbox/S2P Datasets Shared Folder/MLTC_MMC_NH_data/global_cap_reports//nys_medicaid_2024-2025_q2report.pdf")
  
  for(i in c(1:length(report_names))){
    report_name <- report_names[i]
    
    # Extract text from the PDF
    pdf_text <- pdf_text(report_paths[i])
    
    report_table_page <- c(1:length(pdf_text))[pdf_text %>% str_to_lower() %>% str_detect("actual") &
                                                 pdf_text %>% str_to_lower() %>% str_detect("variance") &
                                                 (pdf_text %>% str_to_lower() %>% str_detect("category of service") | 
                                                    pdf_text %>% str_to_lower() %>% str_detect("category of spending") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 1") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 2") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 3") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 4")) &
                                                 pdf_text %>% str_detect("TOTAL") &
                                                 pdf_text %>% str_to_lower() %>% str_detect("table of contents",negate = TRUE)] %>% 
      min()
    
    pdf_table <- pdf_text[report_table_page]
    
    dollar_multiplier <- case_when(pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("in millions") ~ 1000000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("in thousands") ~ 1000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("millions") ~ 1000000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("thousands") ~ 1000,
                                   TRUE ~ 1000000)
    
    # Parse the table from the extracted text
    table_data <- str_split(pdf_table,"\n")[[1]] %>% 
      # Clean and structure the data
      str_trim() %>%
      as_tibble()
    
    table_start_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_to_lower() %>% str_sub(1,19) == "category of service" |
               value %>% str_to_lower() %>% str_sub(1,20) == "category of spending" |
               (value %>% str_to_lower() %>% str_detect("estimated") &
                  value %>% str_to_lower() %>% str_detect("actual") &
                  value %>% str_to_lower() %>% str_detect("variance"))) %>% 
      dplyr::pull(rowid)
    
    first_data_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_count("\\$") == 3,
             rowid > table_start_row) %>% 
      dplyr::pull(rowid) %>%
      min()
    
    table_end_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_sub(1,5) == "TOTAL") %>% 
      dplyr::pull(rowid)
    
    header_rows <- table_data %>% 
      slice(table_start_row:first_data_row - 1) %>% 
      as.character() %>% 
      paste(collapse = " ") %>% 
      str_to_lower()
    
    table_col_names <- case_when(str_detect(header_rows,"target") ~ c("category_of_service","target","actual","variance"),
                                 str_detect(header_rows,"estimated") ~ c("category_of_service","estimated","actual","variance"))
    
    table_clean <- table_data %>%
      slice(first_data_row:table_end_row) %>% 
      separate(col = value, into = c("Column1", "Column2", "Column3", "Column4"), sep = "\\s{2,}", fill = "right")
    
    names(table_clean) <- table_col_names
    
    table_clean <- table_clean %>% 
      mutate(report = report_name %>% str_remove(".pdf"),
             multiplier = dollar_multiplier)
    
    if(i == 1){
      nys_medicaid_spending <- table_clean
    } else{
      nys_medicaid_spending <- bind_rows(nys_medicaid_spending,table_clean)
    }
    paste0(i," of ",length(report_names),"(",scales::percent(i/length(report_names),accuracy = 0.1),") | ",report_name) %>% print()
  }
  write_csv(nys_medicaid_spending,paste0(data_dir,'mmc_mltc/data/Medicaid-Spending-by-Fiscal_Year.csv'))
} else{
  nys_medicaid_spending <- read_csv(paste0(data_dir,'mmc_mltc/data/Medicaid-Spending-by-Fiscal_Year.csv'))
}
```

```{r setup_medicaid_spending_analysis, include=FALSE}
mmc_global_cap_analysis <- nys_medicaid_spending %>% 
  mutate(reference = ifelse(is.na(target),estimated,target),
         ref_type = ifelse(is.na(target),"estimated","target"),
         across(.cols = c("reference","actual"),
                .fns = ~ (.x %>%
                            # Remove $ and ,
                            str_remove_all("[\\$,]") %>%
                            # Replace parentheses with a negative sign
                            str_replace_all("\\((.*)\\)", "-\\1") %>%
                            # Convert to numeric
                            as.numeric()) * multiplier),
         difference = actual - reference) %>% 
  select(category_of_service,
         ref_type,
         reference,
         actual,
         report) %>% 
  # Only include year-end reports in the analysis
  filter(str_detect(report,"q4|Mar")) %>% 
  mutate(fiscal_yr = str_sub(report,14,22),
         type = case_when(category_of_service %in% c("Long Term Managed Care",
                                                     "Nursing Homes",
                                                     "Personal Care",
                                                     "Other Long Term Care") ~ "Long Term Care")) %>% 
  select(-report)
```

```{r setup_nh_analysis, include=FALSE}
nh_trend_data <- nh_historic %>%
  # Step 1: Narrow to relevant columns
  select(facility_id, census_date,
         ends_with("_beds"), ends_with("beds_available"),
         -pediatric_ventilator_beds, -dialysis_stations_beds) %>%
  
  # Step 2: Convert date & derive month/year once
  mutate(census_date = as.Date(census_date, "%m/%d/%Y"),
         census_month = floor_date(census_date, "month"),
         census_year = floor_date(census_date, "year")) %>%
  
  # Step 3: Reshape and clean name/type
  pivot_longer(
    cols = -c(facility_id, census_date, census_month, census_year),
    names_to = "name", values_to = "value"
  ) %>%
  mutate(type = case_when(
    str_ends(name, "beds") ~ "beds",
    str_ends(name, "beds_available") ~ "beds_available"
  ),
  name = name %>%
    str_remove("_beds") %>%
    str_replace_all("_", " ") %>%
    str_to_title()) %>%
  
  # Step 4: Aggregate at facility-date-type level
  group_by(facility_id, census_date, census_month, census_year, type) %>%
  summarise(value = sum(value, na.rm = TRUE), .groups = "drop") %>%
  
  # Step 5: Wide format & calculate census
  pivot_wider(names_from = type, values_from = value) %>%
  mutate(census = beds - beds_available) %>%
  
  # Step 6: Aggregate to state-week level
  group_by(census_year, census_month, census_date) %>%
  summarise(census = sum(census, na.rm = TRUE), .groups = "drop") %>%
  
  # Step 7: Add fiscal year
  mutate(fiscal_yr = if_else(month(census_month) <= 3,
                             year(census_year) - 1,
                             year(census_year)))

nh_monthly_census_graph <- nh_trend_data %>% 
  # Aggregate from State-week level to State-month level (Average)
  group_by(census_month) %>% 
  summarise(census = mean(census)) %>% 
  ungroup() %>% 
  ggplot(aes(x = census_month,
             y = census)) + 
  geom_line() +
  geom_point(size = 1,
             alpha = 0.5) +
  labs(title = "Nursing Home, Monthly Average Census",
       subtitle = paste0("New York State (",
                         # format(min(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                         # " to ",
                         # format(max(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                         "Since ",
                         format(min(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                         ")"),
       x = "",
       y = "Census\n",
       caption = paste0("Data Source: Health Data NY\nData as of: ",format(Sys.Date(),"%B %d, %Y"))) +
  scale_x_date(breaks = breaks_pretty(n = 8)) +
  scale_y_continuous(breaks = breaks_pretty(n = 8),
                     labels = comma_format(accuracy = 1),
                     limits = c(0,NA)) +
  theme_light(base_size = 16) +
  isaac_step2_graph_theme

nh_trend_table_data <- nh_trend_data %>% 
  # Aggregate from State-week level to State-year level (Average)
  group_by(census_year) %>% 
  summarise(census = mean(census)) %>% 
  ungroup() %>% 
  arrange(desc(census_year)) %>% 
  mutate(delta = census - lead(census,n=1),
         delta_pct = delta / lead(census,n=1))

nh_annual_census_table <- nh_trend_table_data %>% 
  gt() %>% 
  tab_header(title = md("**Nursing Home, Annual Average Census**"),
             subtitle = paste0("New York State (",
                               # format(min(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                               # " to ",
                               # format(max(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                               "Since ",
                               format(min(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                               ")")) %>% 
  fmt_date(columns = census_year,
           date_style = "year") %>% 
  fmt_integer(c(census,delta)) %>%
  fmt_percent(delta_pct) %>%
  cols_label(census_year = "Year",
             census = "Mean Census",
             delta = "Increase/Decrease",
             delta_pct = "% Increase/Decrease") %>% 
  cols_align(align = "left", columns = census_year) %>% 
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) %>% 
  tab_source_note(source_note = md("**Data Source:** Health Data NY")) %>% 
  tab_source_note(source_note = md(paste0("**Data as of:** ",format(Sys.Date(),"%B %d, %Y")))) %>% 
  # cols_width(census_year ~ px(175),
  #            all_of(c("census")) ~ px(150)) %>% 
  tab_options(table.font.size = px(10),
              table.font.names = "Helvetica",
              data_row.padding = px(2))

nh_census_vs_spending_table_data <- nh_trend_data %>% 
  group_by(fiscal_yr) %>% 
  summarise(weeks_reported = census_date %>% unique() %>% length(),
            mean_census = mean(census)) %>% 
  ungroup() %>% 
  left_join(nh_trend_data %>% 
              group_by(fiscal_yr) %>% 
              slice_max(census_date) %>% 
              ungroup() %>% 
              select(fiscal_yr,
                     census) %>% 
              rename("end_of_period_census" = "census"),
            by = "fiscal_yr") %>% 
  left_join(mmc_global_cap_analysis %>% 
              mutate(fiscal_yr_dt = fiscal_yr %>% str_sub(1,4) %>% paste("01","01",sep="-") %>% as.Date(),
                     fiscal_yr = fiscal_yr_dt %>% format("%Y") %>% as.numeric) %>% 
              filter(category_of_service == "Nursing Homes") %>% 
              select(fiscal_yr,actual) %>% 
              rename("state_spending" = "actual") %>% 
              mutate(spending = state_spending * 2) %>% 
              select(-state_spending),
            by = "fiscal_yr") %>% 
  
  # left_join(nh_trend_data %>% 
  #             arrange(desc(census_date)) %>% 
  #             mutate(census_delta = census  - lead(census),
  #                    census_delta_pct = census_delta / lead(census,n=1)) %>% 
  #             group_by(fiscal_yr) %>% 
  #             summarise(mean_weekly_change_pct = mean(census_delta_pct)) %>% 
  #             ungroup(),
  #           by = "fiscal_yr") %>% 
  
  mutate(spending_per_capita = spending / mean_census) %>% 
  arrange(desc(fiscal_yr)) %>% 
  filter(fiscal_yr >= 2011) %>% 
  mutate(across(.cols = c("mean_census","spending","spending_per_capita"),
                .fns = ~ .x  - lead(.x),
                .names = "{.col}_delta"),
         mean_census_delta_pct = mean_census_delta / lead(mean_census,n=1),
         spending_delta_pct = spending_delta / lead(spending,n=1),
         spending_per_capita_delta_pct = spending_per_capita_delta / lead(spending_per_capita,n=1),
         fiscal_yr = paste(fiscal_yr,str_sub(fiscal_yr + 1,3,4),sep = "-"),
         fiscal_yr = paste0("FY",str_sub(fiscal_yr,-2))) %>% 
  relocate(fiscal_yr,
           weeks_reported,
           mean_census,
           mean_census_delta,
           mean_census_delta_pct,
           end_of_period_census,
           spending,
           spending_delta,
           spending_delta_pct,
           spending_per_capita,
           spending_per_capita_delta,
           spending_per_capita_delta_pct)

nh_census_vs_spending_table <- nh_census_vs_spending_table_data %>% 
  gt() %>% 
  tab_header(title = md("**Nursing Home Census and Spending, by Fiscal Year**"),
             subtitle = paste0("New York State (",
                               # format(min(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                               # " to ",
                               # format(max(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                               "Since ",
                               format(min(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                               ")")) %>%
  fmt_integer(c(weeks_reported,
                mean_census,
                end_of_period_census,
                mean_census_delta)) %>% 
  fmt_currency(c(spending,
                 spending_delta,
                 spending_per_capita,
                 spending_per_capita_delta),
               decimals = 0) %>% 
  fmt_percent(ends_with("pct")) %>% 
  tab_spanner(label = "Census",
              columns = c(mean_census,
                          mean_census_delta,
                          mean_census_delta_pct,
                          end_of_period_census)) %>% 
  tab_spanner(label = "Gross Spending",
              columns = c(spending,
                          spending_delta,
                          spending_delta_pct)) %>% 
  tab_spanner(label = "Average Annual Gross Spending per Capita",
              columns = c(spending_per_capita,
                          spending_per_capita_delta,
                          spending_per_capita_delta_pct)) %>% 
  cols_label(fiscal_yr = "Fiscal Year",
             weeks_reported = "Weeks Reported",
             mean_census = "Average Census",
             mean_census_delta = "Annual Change",
             mean_census_delta_pct = "Annual Change %",
             end_of_period_census = "End of Period Census",
             spending = "Medicaid Fee-for-Service Spending",
             spending_delta = "Annual Change",
             spending_delta_pct = "Annual Change %",
             spending_per_capita = "Medicaid Fee-for-Service Spending",
             spending_per_capita_delta = "Annual Change",
             spending_per_capita_delta_pct = "Annual Change %") %>% 
  cols_align(align = "left", columns = fiscal_yr) %>% 
  tab_style(style = cell_text(weight = "bold"),
            locations = list(cells_column_labels(),
                             cells_column_spanners())) %>% 
  tab_source_note(source_note = md("**Data Sources:** Health Data NY; NYSDOH Global Spending Cap
Reports")) %>% 
  tab_source_note(source_note = md(paste0("**Data as of:** ",format(Sys.Date(),"%B %d, %Y")))) %>% 
  tab_options(table.font.size = px(10),
              table.font.names = "Helvetica",
              data_row.padding = px(2))

# Calculate scaling factors for normalization
max_census <- max(nh_census_vs_spending_table_data$mean_census,na.rm=TRUE)
max_spending <- max(nh_census_vs_spending_table_data$spending,na.rm=TRUE) / 1e9

nh_census_vs_spending_graph <- nh_census_vs_spending_table_data %>% 
  select(fiscal_yr,
         mean_census,
         spending) %>% 
  # Create the plot
  ggplot(aes(x = fiscal_yr)) +
  # Plot enrollee_months as bars
  geom_bar(aes(y = mean_census, fill = "Census"),
           stat = "identity", alpha = 0.9) +
  # Plot mltc_spending as a line with points
  geom_line(aes(y = (spending / 1e9) * (max_census / max_spending),
                group = 1, color = "Spending")) +
  geom_point(aes(y = (spending / 1e9) * (max_census / max_spending)), color = "black") +
  # Left y-axis for mean_census
  scale_y_continuous(
    name = "Average Census\n",
    labels = scales::label_comma(),
    breaks = breaks_pretty(n = 8),
    sec.axis = sec_axis(~ . * (max_spending / max_census), 
                        name = "Medicaid Fee-For-Service Gross Spending\n", 
                        breaks = breaks_pretty(n = 8),
                        labels = scales::label_dollar(scale = 1,
                                                      accuracy = 1,
                                                      suffix = "B"))) +
  scale_fill_manual(values = "#B9CDE5") +
  scale_color_manual(values = "black") +
  # Formatting
  labs(x = "\nFiscal Year",
       title = "Nursing Home Census and Spending, by Fiscal Year",
       subtitle = paste0("New York State (",
                         # format(min(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                         # " to ",
                         # format(max(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                         "Since ",
                         format(min(as.Date(nh_historic$census_date,"%m/%d/%Y")),"%b %Y"),
                         ")"),
       color = "",
       fill = "",
       caption = paste0("Data Sources: Health Data NY; NYSDOH Global Spending Cap Reports\nData as of: ",format(Sys.Date(),"%B %d, %Y"))) +
  theme_light(base_size = 16) +
  isaac_step2_graph_theme +
  theme(legend.text = element_text(face = "bold"))
```

<br>

### Census and Spending

<br>

```{r nh_enrollment_vs_spending_table}
nh_census_vs_spending_table %>% 
  tab_source_note(
    nh_census_vs_spending_table_data %>% 
      download_this(
        output_name = paste0("nh_census_and_spending_",str_replace_all(Sys.Date(),"-","_")),
        output_extension = ".xlsx",
        button_label = "Download Excel Workbook",
        button_type = "default",
        class = "button_download"
      )
  )
```
**Data Sources:** [Health Data NY](https://health.data.ny.gov/Health/Nursing-Home-Weekly-Bed-Census-Beginning-2009/uhyy-xp9s/about_data); [New York State Department of Health](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/)

<br>

```{r nh_census_vs_spending_graph}
nh_census_vs_spending_graph
```

**Data Sources:** [Health Data NY](https://health.data.ny.gov/Health/Nursing-Home-Weekly-Bed-Census-Beginning-2009/uhyy-xp9s/about_data); [New York State Department of Health](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/)

<br>

### Census, Trends Over Time

<br>

```{r nh_annual_census_table}
nh_annual_census_table
```

**Data Source:** [Health Data NY](https://health.data.ny.gov/Health/Nursing-Home-Weekly-Bed-Census-Beginning-2009/uhyy-xp9s/about_data)

<br>

```{r nh_monthly_census_graph}
nh_monthly_census_graph
```

**Data Source:** [Health Data NY](https://health.data.ny.gov/Health/Nursing-Home-Weekly-Bed-Census-Beginning-2009/uhyy-xp9s/about_data)

<br>

***

### Technical Notes

1. Data include nursing home beds, pediatric beds, behavioral intervention beds, ventilator beds, scatter ventilator beds, traumatic brain injury beds, neurodegenerative disease beds; exclude pediatric ventilator beds, dialysis station beds, and adult day health care slots.  
2. Spending data represent Medicaid fee-for-service spending only. Spending may include other minor related categories of service spending, such as adult day health care.
3. Gross spending is estimated assuming a [50% federal match](https://www.kff.org/medicaid/state-indicator/federal-matching-rate-and-multiplier/?currentTimeframe=0&sortModel=%7B%22colId%22:%22Location%22,%22sort%22:%22asc%22%7D) to the State spending reported in quarterly Global Cap Reports.
4. Census data represent all occupied beds, regardless of payor type.
5. <i>Beds</i> = The number of Residential Beds the facility is approved to operate.  
6. <i>Beds Available</i> = Available Residential Beds reported by facility. This figure should include any licensed beds that are not operational.  
7. <i>Census (aka "Occupied Beds")</i> = (Beds) - (Beds Available).  
8. <i>End of Period Census</i> The total number of occupied beds at the end of the reporting period. For completed fiscal years, this represents the census at the end of the year. For partial fiscal years, it reflects the census as of the most recent reporting date.												
