---
output:
  html_document:
    toc: false
params:
  project_directory: "/Users/isaac/Dropbox/S2P Datasets Shared Folder/MLTC_MMC_NH_data/"
  use_local_data: FALSE
---

<style>
.button_download {
font-size: 16px;
font-family: "Arial";
color: black;
background-color: #C7C9D8;
}
</style>

```{r setup_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, error = TRUE, fig.width = 12, fig.height = 8, cache = FALSE)
```

```{r setup_libraries, include=FALSE}
library(tidyverse)
library(furrr)
library(scales)
library(gt)
library(pdftools)
library(httr)
library(downloadthis)
plan(multisession)
data_dir <- params$project_directory
```

```{r isaac_step2_graph_theme, include=FALSE}
isaac_step2_graph_theme <- theme(plot.title.position = "plot",
                                 plot.title = element_text(face = "bold"),
                                 legend.position = "top",
                                 strip.text = element_text(color = "black"),
                                 legend.title = element_text(face = "bold",hjust = 0.5),
                                 axis.title.y = element_text(face = "bold",angle = 90),
                                 axis.title.x = element_text(face = "bold"))
```

```{r setup_medicaid_enrollment_data_ingestion, include=FALSE}
if(params$use_local_data == FALSE){
  # # https://health.data.ny.gov/Health/Medicaid-Program-Enrollment-by-Month-Beginning-200/m4hz-kzn3/about_data
  # nys_medicaid <- read_csv("https://health.data.ny.gov/api/views/m4hz-kzn3/rows.csv?accessType=DOWNLOAD") %>% janitor::clean_names()
  
  # https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/
  # https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/archives.htm
  
  # Custom headers for the request
  custom_headers <- add_headers(
    `User-Agent` = "Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/97.0.4692.71 Safari/537.36",
    `Accept` = "application/vnd.openxmlformats-officedocument.spreadsheetml.sheet, application/vnd.ms-excel;q=0.9,*/*;q=0.8",
    `Connection` = "keep-alive"
  )
  
  # Function to download a file
  download_file <- function(url, save_directory, filename) {
    destfile <- file.path(save_directory, filename)  # Create full path
    response <- GET(url, custom_headers)
    
    if (status_code(response) == 200) {
      writeBin(content(response, "raw"), destfile)
      message("File saved to: ", destfile)
    } else {
      warning("Failed to download file: ", url, " - Status code: ", status_code(response))
    }
  }
  
  # Define months and years
  months <- sprintf("%02d", 1:12)  # "01", "02", ..., "12"
  years <- 2008:as.numeric(format(Sys.Date(),"%Y"))              # Adjust range as needed
  
  already_downloaded_mmc_files <- list.files(paste0(data_dir,"mmc_mltc/data"),recursive = TRUE)
  
  # Loop through years and months to construct URLs and download files
  for (year in years) {
    for (month in months) {
      
      if(paste0("en",month,"_",year,".xls") %in% already_downloaded_mmc_files | paste0("en",month,"_",year,".xlsx") %in% already_downloaded_mmc_files){
        message(paste0("en",month,"_",year," was already downloaded"))
      } else{
        
        # Construct the URL for both file types
        url_xlsx <- paste0("https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/", year, "/docs/en", month, "_", substr(year, 3, 4), ".xlsx")
        url_xls <- paste0("https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/", year, "/docs/en", month, "_", substr(year, 3, 4), ".xls")
        
        # Attempt to download .xlsx file first
        tryCatch(
          download_file(url_xlsx, paste0(data_dir,"mmc_mltc/data"), paste0("en", month, "_", year, ".xlsx")),
          error = function(e) {
            message("Trying .xls for ", month, "/", year)
          }
        )
        
        # Attempt to download .xls file if .xlsx fails
        tryCatch(
          download_file(url_xls, paste0(data_dir,"mmc_mltc/data"), paste0("en", month, "_", year, ".xls")),
          error = function(e) {
            message("Failed for both .xlsx and .xls for ", month, "/", year)
          }
        )
      }
    }
  }
  
  mmc_county_list <- c("ALBANY","ALLEGANY","BROOME","CATTARAUGUS","CAYUGA","CHAUTAUQUA","CHEMUMG","CHEMUNG","CHENANGO",
                       "CLINTON","COLUMBIA","CORTLAND","DELAWARE","DUTCHESS","ERIE","ESSEX","FRANKLIN","FULTON",
                       "GENESEE","GREENE","HAMILTON","HERKIMER","JEFFERSON","LEWIS","LIVINGSTON","MADISON","MONROE",
                       "MONTGOMERY","NASSAU","NEW YORK","NEW YORK CITY","NIAGARA","ONEIDA","ONONDAGA","ONTARIO","ORANGE",
                       "ORLEANS","OSWEGO","OTSEGO","PUTNAM","RENSSELAER","ROCKLAND","SAINT LAWRENCE","SARATOGA","SCHENECTADY", 
                       "SCHOHARIE","SCHUYLER","SENECA","ST LAWRENCE","STEUBEN","SUFFOLK","SULLIVAN","TIOGA","TOMPKINS",
                       "ULSTER","WARREN","WASHINGTON","WAYNE","WESTCHESTER","WYOMING","YATES")
  
  downloaded_mmc_files <- list.files(paste0(data_dir,"mmc_mltc/data"),
                                     recursive = TRUE,
                                     full.names = TRUE,
                                     pattern = "xls")
  
  report_sheets <- downloaded_mmc_files %>% 
    str_replace("en10_24","en10_2024") %>% 
    tibble() %>% 
    rename("report" = 1) %>% 
    filter(str_detect(report,"Medicaid-Program-Enrollment-by-Month-Beginning-2009.csv",negate = TRUE),
           str_detect(report,"Medicaid-Spending-by-Fiscal_Year.csv",negate = TRUE)) %>% 
    rowwise() %>% 
    mutate(sheets = readxl::excel_sheets(report) %>% paste(collapse = ";")) %>% 
    ungroup() %>% 
    separate_rows(sheets,sep = ";") %>% 
    mutate(report_my = report %>%
             str_remove(paste0(data_dir,"mmc_mltc/data/")) %>% 
             str_remove(".xlsx") %>% 
             str_remove(".xls") %>% 
             str_remove("en") %>%
             paste0(.,"_01") %>% 
             as.Date("%m_%Y_%d"))
  
  parse_mmc <- function(input_file){
    
    mmc_tab_name <- readxl::excel_sheets(input_file)
    mmc_tab_name <- mmc_tab_name[str_detect(mmc_tab_name,"Medicaid Managed")]
    
    temp_1 <- readxl::read_excel(input_file,
                               sheet = mmc_tab_name,
                               skip = 0) %>%
      janitor::clean_names()
    
    temp_2 <- temp_1 %>% 
      rename("col_1" = 1,
             "col_2" = 2) %>% 
      filter(str_detect(str_to_lower(col_1),"total") | str_detect(str_to_lower(col_2),"total"),
             str_detect(str_to_lower(col_1),"state|grand") | str_detect(str_to_lower(col_2),"state|grand"),
             str_detect(str_to_lower(col_1),"upstate",negate = TRUE) | is.na(col_1),
             str_detect(str_to_lower(col_2),"upstate",negate = TRUE) | is.na(col_2))
    
    if(nrow(temp_2) == 0){
      temp_1 %>% 
        select(ncol(temp_1)) %>% 
        rename("col_1" = 1) %>% 
        mutate(col_1 = col_1 %>% 
                 as.character() %>% 
                 str_remove_all(",") %>% 
                 as.numeric()) %>% 
        summarise(number_of_recipients = max(col_1,na.rm = TRUE)) %>% 
        return()
        
    } else{
      temp_2 %>% 
        slice(1) %>% 
        select(-c(1:2)) %>% 
        pivot_longer(cols = everything(),
                     values_to = "number_of_recipients",
                     values_transform = as.character) %>% 
        mutate(number_of_recipients = number_of_recipients %>% 
                 str_remove_all(",") %>% 
                 as.numeric()) %>% 
        summarise(number_of_recipients = max(number_of_recipients,na.rm = TRUE)) %>% 
        return()
    }
  }
  
  start_time_mmc <- Sys.time()
  mmc <- report_sheets %>%
    filter(sheets %>% str_to_lower() %>% str_detect("medicaid managed"),
           report_my >= as.Date("2008-08-01")) %>%
    future_map2_dfr(.x = dplyr::pull(.,report),
                    .y = dplyr::pull(.,report_my),
                    .f = ~ parse_mmc(.x) %>%
                      mutate(month_year = .y),
                    .progress = TRUE) %>% 
    relocate(month_year)
  end_time_mmc <- Sys.time()
  # end_time_mmc - start_time_mmc
  
  nys_medicaid <- mmc %>% 
    mutate(eligibility_year = format(month_year,"%Y") %>% as.numeric,
           eligibility_month = format(month_year,"%m") %>% as.numeric) %>% 
    select(eligibility_year,
           eligibility_month,
           number_of_recipients)
  
  write_csv(nys_medicaid,paste0(data_dir,'mmc_mltc/data/nys_mmc_enrollment.csv'))
} else{
  nys_medicaid <- read_csv(paste0(data_dir,'mmc_mltc/data/nys_mmc_enrollment.csv'))
}
```

```{r setup_medicaid_spending_data_ingestion, include=FALSE}
if(params$use_local_data == FALSE){
  # https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/
  try_downloading_report <- function(){
    # Use GET with a User-Agent header
    response <- GET(pdf_url, user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/121.0.0.0 Safari/537.36"))
    
    # Check response status
    if (status_code(response) == 200) {
      writeBin(content(response, "raw"), pdf_file)
      message("File downloaded successfully.")
    } else {
      message("Failed to download file. HTTP status: ", status_code(response))
    }
  }
  
  # Download Global Cap Quarterly Reports
  for(i in c(2011:(as.numeric(format(Sys.Date(),"%Y"))))){
    for(j in c(1:4)){
      qrt_number <- case_when(j == 1 ~ "1st",
                              j == 2 ~ "2nd",
                              j == 3 ~ "3rd",
                              j == 4 ~ "4th")
      
      pdf_url <- paste0("https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/monthly/sfy_",i,"-",i+1,"/docs/",qrt_number,"_qtr_rpt.pdf")
      
      pdf_file <- paste0(data_dir,"global_cap_reports/nys_medicaid_",i,"-",i+1,"_q",j,"report.pdf")
      
      if(file.exists(pdf_file) == FALSE){
        try_downloading_report()
      } else{
        message("File already downloaded")
      }
    }
  }
  
  # Download Global Cap Monthly Reports
  for(i in c(2011:(as.numeric(format(Sys.Date(),"%Y"))))){
    for(j in c(month.name,month.abb,"Sept")){
      for(k in c(0,1)){
        pdf_url <- paste0("https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/monthly/sfy_",i,"-",i+1,"/docs/",j,"_",i+k,"_report.pdf") %>% 
          str_to_lower()
        
        pdf_file <- paste0(data_dir,"global_cap_reports/nys_medicaid_",i,"-",i+1,"_month_",j,"_",i+k,"_report.pdf")
        
        if(file.exists(pdf_file) == FALSE){
          try_downloading_report()
        } else{
          message("File already downloaded")
        }
      }
    }
  }
  
  report_names <- list.files(paste0(data_dir,"global_cap_reports/"),
                             pattern = ".pdf",
                             full.names = FALSE,
                             recursive = FALSE) %>% 
    # Unusually formatted non-Q4 report
    setdiff("nys_medicaid_2024-2025_q2report.pdf")
  
  report_paths <- list.files(paste0(data_dir,"global_cap_reports/"),
                             pattern = ".pdf",
                             full.names = TRUE,
                             recursive = FALSE) %>% 
    # Unusually formatted non-Q4 report
    setdiff("/Users/isaac/Dropbox/S2P Datasets Shared Folder/MLTC_MMC_NH_data/global_cap_reports//nys_medicaid_2024-2025_q2report.pdf")
  
  for(i in c(1:length(report_names))){
    report_name <- report_names[i]
    
    # Extract text from the PDF
    pdf_text <- pdf_text(report_paths[i])
    
    report_table_page <- c(1:length(pdf_text))[pdf_text %>% str_to_lower() %>% str_detect("actual") &
                                                 pdf_text %>% str_to_lower() %>% str_detect("variance") &
                                                 (pdf_text %>% str_to_lower() %>% str_detect("category of service") | 
                                                    pdf_text %>% str_to_lower() %>% str_detect("category of spending") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 1") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 2") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 3") |
                                                    pdf_text %>% str_to_lower() %>% str_detect("quarter 4")) &
                                                 pdf_text %>% str_detect("TOTAL") &
                                                 pdf_text %>% str_to_lower() %>% str_detect("table of contents",negate = TRUE)] %>% 
      min()
    
    pdf_table <- pdf_text[report_table_page]
    
    dollar_multiplier <- case_when(pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("in millions") ~ 1000000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("in thousands") ~ 1000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("millions") ~ 1000000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("thousands") ~ 1000,
                                   TRUE ~ 1000000)
    
    # Parse the table from the extracted text
    table_data <- str_split(pdf_table,"\n")[[1]] %>% 
      # Clean and structure the data
      str_trim() %>%
      as_tibble()
    
    table_start_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_to_lower() %>% str_sub(1,19) == "category of service" |
               value %>% str_to_lower() %>% str_sub(1,20) == "category of spending" |
               (value %>% str_to_lower() %>% str_detect("estimated") &
                  value %>% str_to_lower() %>% str_detect("actual") &
                  value %>% str_to_lower() %>% str_detect("variance"))) %>% 
      dplyr::pull(rowid)
    
    first_data_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_count("\\$") == 3,
             rowid > table_start_row) %>% 
      dplyr::pull(rowid) %>%
      min()
    
    table_end_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_sub(1,5) == "TOTAL") %>% 
      dplyr::pull(rowid)
    
    header_rows <- table_data %>% 
      slice(table_start_row:first_data_row - 1) %>% 
      as.character() %>% 
      paste(collapse = " ") %>% 
      str_to_lower()
    
    table_col_names <- case_when(str_detect(header_rows,"target") ~ c("category_of_service","target","actual","variance"),
                                 str_detect(header_rows,"estimated") ~ c("category_of_service","estimated","actual","variance"))
    
    table_clean <- table_data %>%
      slice(first_data_row:table_end_row) %>% 
      separate(col = value, into = c("Column1", "Column2", "Column3", "Column4"), sep = "\\s{2,}", fill = "right")
    
    names(table_clean) <- table_col_names
    
    table_clean <- table_clean %>% 
      mutate(report = report_name %>% str_remove(".pdf"),
             multiplier = dollar_multiplier)
    
    if(i == 1){
      nys_medicaid_spending <- table_clean
    } else{
      nys_medicaid_spending <- bind_rows(nys_medicaid_spending,table_clean)
    }
    paste0(i," of ",length(report_names),"(",scales::percent(i/length(report_names),accuracy = 0.1),") | ",report_name) %>% print()
  }
  write_csv(nys_medicaid_spending,paste0(data_dir,'mmc_mltc/data/Medicaid-Spending-by-Fiscal_Year.csv'))
} else{
  nys_medicaid_spending <- read_csv(paste0(data_dir,'mmc_mltc/data/Medicaid-Spending-by-Fiscal_Year.csv'))
}
```

```{r setup_medicaid_enrollment_vs_spending_analysis, include=FALSE}
mmc_analysis <- nys_medicaid %>% 
  # filter(managed_care_vs_fee_for_service == "MMC") %>% 
  mutate(month_year = paste(eligibility_year,eligibility_month,"01",sep = "-") %>% as.Date(),
         year = paste(eligibility_year,"01","01",sep = "-") %>% as.Date()
         # ,
         # plan_type = case_when(plan_type == "PARTIAL CAP PLAN" ~ "PARTIAL CAP PLAN and PARTIAL MLTC",
         #                       plan_type == "PARTIAL MLTC" ~ "PARTIAL CAP PLAN and PARTIAL MLTC",
         #                       TRUE ~ plan_type)
  )

mmc_global_cap_analysis <- nys_medicaid_spending %>% 
  mutate(reference = ifelse(is.na(target),estimated,target),
         ref_type = ifelse(is.na(target),"estimated","target"),
         across(.cols = c("reference","actual"),
                .fns = ~ (.x %>%
                            # Remove $ and ,
                            str_remove_all("[\\$,]") %>%
                            # Replace parentheses with a negative sign
                            str_replace_all("\\((.*)\\)", "-\\1") %>%
                            # Convert to numeric
                            as.numeric()) * multiplier),
         difference = actual - reference) %>% 
  select(category_of_service,
         ref_type,
         reference,
         actual,
         report) %>% 
  # Only include year-end reports in the analysis
  filter(str_detect(report,"q4|Mar")) %>% 
  mutate(fiscal_yr = str_sub(report,14,22),
         type = case_when(category_of_service %in% c("Mainstream Managed Care",
                                                     "Inpatient",
                                                     "Outpatient/Emergency Room",
                                                     "Clinic",
                                                     "Pharmacy",
                                                     "Transportation",
                                                     "Non-Institutional",
                                                     "Non-Institutional / Other") ~ "Mainstream",
                          category_of_service %in% c("Other Long Term Care") ~ "Long Term Care")) %>% 
  select(-report)

mmc_global_cap_analysis_data_v1 <- mmc_global_cap_analysis %>% 
  mutate(fiscal_yr_dt = fiscal_yr %>% str_sub(1,4) %>% paste("01","01",sep="-") %>% as.Date()) %>% 
  select(-ref_type,
         -reference) %>% 
  filter(category_of_service %in% c("Medicaid Managed Care",
                                    "Managed Care",
                                    "Total Fee-For-Service",
                                    "Total Fee For Service")) %>% 
  mutate(category_of_service_type = case_when(category_of_service %>% str_to_lower() %>% str_detect("fee") ~ "Fee For Service",
                                              category_of_service %>% str_to_lower() %>% str_detect("managed") ~ "Managed Care"))

mmc_enrollment_vs_spending_table_data <- mmc_analysis %>% 
  mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                               TRUE ~ eligibility_year)) %>% 
  group_by(fiscal_yr) %>% 
  summarise(enrollee_months = sum(number_of_recipients)) %>% 
  ungroup() %>% 
  left_join(mmc_analysis %>% 
              mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                                           TRUE ~ eligibility_year)) %>% 
              group_by(fiscal_yr,eligibility_month) %>% 
              summarise(enrollee_months = sum(number_of_recipients)) %>% 
              ungroup() %>% 
              arrange(desc(fiscal_yr),
                      desc(eligibility_month)) %>% 
              mutate(enrollee_months_delta = enrollee_months  - lead(enrollee_months),
                     enrollee_months_delta_pct = enrollee_months_delta / lead(enrollee_months,n=1)) %>% 
              group_by(fiscal_yr) %>% 
              summarise(mean_monthly_change_pct = mean(enrollee_months_delta_pct)) %>% 
              ungroup(),
            by = "fiscal_yr") %>% 
  left_join(mmc_analysis %>% 
              mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                                           TRUE ~ eligibility_year)) %>% 
              group_by(fiscal_yr,eligibility_month) %>% 
              summarise(enrollee_months = sum(number_of_recipients)) %>% 
              ungroup() %>% 
              group_by(fiscal_yr) %>% 
              slice_max(eligibility_month) %>% 
              ungroup() %>% 
              select(fiscal_yr,
                     enrollee_months) %>% 
              rename("end_of_period_enrollment" = "enrollee_months"),
            by = "fiscal_yr") %>% 
  arrange(desc(fiscal_yr)) %>% 
  left_join(mmc_global_cap_analysis %>% 
              mutate(fiscal_yr_dt = fiscal_yr %>% str_sub(1,4) %>% paste("01","01",sep="-") %>% as.Date(),
                     fiscal_yr = fiscal_yr_dt %>% format("%Y") %>% as.numeric) %>% 
              filter(category_of_service %in% c("Medicaid Managed Care",
                                                "Managed Care")) %>% 
              select(fiscal_yr,actual) %>% 
              rename("mmc_spending" = "actual"),
            by = "fiscal_yr") %>% 
  left_join(mmc_analysis %>% 
              mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                                           TRUE ~ eligibility_year)) %>% 
              group_by(fiscal_yr,eligibility_month) %>% 
              summarise(enrollee_months = sum(number_of_recipients)) %>% 
              ungroup() %>% 
              group_by(fiscal_yr) %>% 
              summarise(n_months = n(),
                        mean_number_of_enrollees = mean(enrollee_months)) %>% 
              ungroup(),
            by = "fiscal_yr") %>% 
  filter(fiscal_yr >= 2011) %>% 
  mutate(mmc_spending_delta = mmc_spending  - lead(mmc_spending),
         mmc_spending_delta_pct = mmc_spending_delta / lead(mmc_spending,n=1),,
         spending_per_enrollee_month = mmc_spending / enrollee_months,
         spending_per_enrollee_month_delta = spending_per_enrollee_month  - lead(spending_per_enrollee_month),
         spending_per_enrollee_month_delta_pct = spending_per_enrollee_month_delta / lead(spending_per_enrollee_month,n=1),
         enrollee_month_delta = enrollee_months - lead(enrollee_months),
         enrollee_month_delta_pct = enrollee_month_delta / lead(enrollee_months,n=1),
         fiscal_yr = paste(fiscal_yr,str_sub(fiscal_yr + 1,3,4),sep = "-"),
         fiscal_yr = paste0("FY",str_sub(fiscal_yr,-2)),
         mean_spend_per_enrollee = mmc_spending / mean_number_of_enrollees,
         mean_spend_per_enrollee_delta = mean_spend_per_enrollee - lead(mean_spend_per_enrollee),
         mean_spend_per_enrollee_delta_pct = mean_spend_per_enrollee_delta / lead(mean_spend_per_enrollee,n=1),
         across(.cols = c(enrollee_month_delta,enrollee_month_delta_pct),
                .fn = ~ ifelse(n_months != 12,NA,.x))) %>% 
  select(fiscal_yr,
         n_months,
         enrollee_months,
         mean_monthly_change_pct,
         enrollee_month_delta,
         enrollee_month_delta_pct,
         end_of_period_enrollment,
         mmc_spending,
         mmc_spending_delta,
         mmc_spending_delta_pct,
         mean_number_of_enrollees,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta,
         mean_spend_per_enrollee_delta_pct)

if(mmc_enrollment_vs_spending_table_data %>% slice(1) %>% dplyr::pull(n_months) < 12){
  projection_data_ytd <- mmc_enrollment_vs_spending_table_data %>% slice(1)
  projection_data_base_year <- mmc_enrollment_vs_spending_table_data %>% slice(2)
  projection_data_base_5_years <- mmc_enrollment_vs_spending_table_data %>% slice(2:6)
  
  fiscal_yr_projected <- paste0(projection_data_ytd$fiscal_yr," (Estimate)")
  
  enrollee_months_projected <- projection_data_ytd$enrollee_months / (projection_data_ytd$n_months / 12)
  end_of_period_enrollment_projected <- enrollee_months_projected / 12
  mean_number_of_enrollees_projected <- ((enrollee_months_projected + projection_data_base_year$enrollee_months) / 2) / 12
  enrollee_month_delta_projected <- enrollee_months_projected - projection_data_base_year$enrollee_months
  enrollee_month_delta_pct_projected <- enrollee_month_delta_projected / projection_data_base_year$enrollee_months
  
  spending_growth_rate_projected <- mean(projection_data_base_5_years$mean_spend_per_enrollee_delta_pct)
  mean_spend_per_enrollee_projected <- (1 + spending_growth_rate_projected) * projection_data_base_year$mean_spend_per_enrollee
  mmc_spending_projected <- mean_number_of_enrollees_projected * mean_spend_per_enrollee_projected
  mmc_spending_delta_projected <- mmc_spending_projected - projection_data_base_year$mmc_spending
  mmc_spending_delta_pct_projected <- mmc_spending_delta_projected / projection_data_base_year$mmc_spending
  mean_spend_per_enrollee_delta_projected <- mean_spend_per_enrollee_projected - projection_data_base_year$mean_spend_per_enrollee
  mean_spend_per_enrollee_delta_pct_projected <- mean_spend_per_enrollee_delta_projected / projection_data_base_year$mean_spend_per_enrollee
  
  mmc_enrollment_vs_spending_projection_table_data <- tibble(
    fiscal_yr = fiscal_yr_projected,
    enrollee_months = enrollee_months_projected,
    enrollee_month_delta = enrollee_month_delta_projected,
    enrollee_month_delta_pct = enrollee_month_delta_pct_projected,
    end_of_period_enrollment = end_of_period_enrollment_projected,
    mmc_spending = mmc_spending_projected,
    mmc_spending_delta = mmc_spending_delta_projected,
    mmc_spending_delta_pct = mmc_spending_delta_pct_projected,
    mean_number_of_enrollees = mean_number_of_enrollees_projected,
    mean_spend_per_enrollee = mean_spend_per_enrollee_projected,
    mean_spend_per_enrollee_delta = mean_spend_per_enrollee_delta_projected,
    mean_spend_per_enrollee_delta_pct = mean_spend_per_enrollee_delta_pct_projected
  )
} else{
  mmc_enrollment_vs_spending_projection_table_data <- tibble()
}

mmc_enrollment_vs_spending_table <- bind_rows(mmc_enrollment_vs_spending_projection_table_data,
                                              mmc_enrollment_vs_spending_table_data %>% 
                                                mutate(fiscal_yr = ifelse(n_months < 12,paste0(fiscal_yr," (YTD)"),fiscal_yr))) %>% 
    relocate(any_of(names(mmc_enrollment_vs_spending_table_data))) %>% 
  gt() %>% 
  tab_header(title = md("**Medicaid Managed Care Enrollment and Spending, by Fiscal Year**"),
             subtitle = paste0("New York State (",
                               # mmc_global_cap_analysis_data_v1 %>% slice_min(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                               # " to ",
                               # mmc_global_cap_analysis_data_v1 %>% slice_max(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                               "Since April 2011",
                               ")")) %>%
  fmt_integer(c(n_months,
                enrollee_months,
                enrollee_month_delta,
                mean_number_of_enrollees,
                end_of_period_enrollment)) %>% 
  fmt_currency(c(mmc_spending,
                 mmc_spending_delta,
                 mean_spend_per_enrollee,
                 mean_spend_per_enrollee_delta),
               decimals = 0) %>% 
  fmt_percent(c(enrollee_month_delta_pct,
                mmc_spending_delta_pct,
                mean_spend_per_enrollee_delta_pct,
                mean_monthly_change_pct)) %>% 
  tab_spanner(label = "Enrollment",
              columns = c(enrollee_months,
                          mean_monthly_change_pct,
                          enrollee_month_delta,
                          enrollee_month_delta_pct,
                          end_of_period_enrollment)) %>% 
  tab_spanner(label = "State Spending",
              columns = c(mmc_spending,
                          mmc_spending_delta,
                          mmc_spending_delta_pct)) %>% 
  tab_spanner(label = "Average Annual Spending per Enrollee",
              columns = c(mean_number_of_enrollees,
                          mean_spend_per_enrollee,
                          mean_spend_per_enrollee_delta,
                          mean_spend_per_enrollee_delta_pct)) %>% 
  cols_label(fiscal_yr = "Fiscal Year",
             n_months = "Months Reported",
             enrollee_months = "Enrollee Months",
             mean_monthly_change_pct = "Average Monthly Change %",
             enrollee_month_delta = "Annual Change",
             enrollee_month_delta_pct = "Annual Change %",
             end_of_period_enrollment = "End of Period Enrollment",
             mmc_spending = "Spending",
             mmc_spending_delta = "Annual Change",
             mmc_spending_delta_pct = "Annual Change %",
             mean_number_of_enrollees = "Average Monthly Number of Enrollees",
             mean_spend_per_enrollee = "Spending",
             mean_spend_per_enrollee_delta = "Annual Change",
             mean_spend_per_enrollee_delta_pct = "Annual Change %") %>% 
  cols_align(align = "left", columns = fiscal_yr) %>% 
  tab_style(style = cell_text(weight = "bold"),
            locations = list(cells_column_labels(),
                             cells_column_spanners())) %>% 
  tab_source_note(source_note = md("**Data Sources:** New York State Department of Health Medicaid Managed Care Enrollment Reports, and Medicaid Global Spending Cap Updates")) %>% 
  tab_source_note(source_note = md(paste0("**Data as of:** ",format(Sys.Date(),"%B %d, %Y")))) %>% 
  tab_options(table.font.size = px(10),
              table.font.names = "Helvetica",
              data_row.padding = px(2),
              table.width = pct(95),
              container.width = px(1000))

# Calculate scaling factors for normalization
max_enrollee_mmc <- max(mmc_enrollment_vs_spending_table_data$enrollee_months,na.rm=TRUE) / 1e6
max_spending_mmc <- max(mmc_enrollment_vs_spending_table_data$mmc_spending,na.rm=TRUE) / 1e9

mmc_enrollment_vs_spending_graph <- mmc_enrollment_vs_spending_table_data %>% 
  select(fiscal_yr,
         enrollee_months,
         mmc_spending) %>% 
  # Create the plot
  ggplot(aes(x = fiscal_yr)) +
  # Plot enrollee_months as bars
  geom_bar(aes(y = enrollee_months / 1e6, fill = "Enrollment"),
           stat = "identity", alpha = 0.9) +
  # Plot mltc_spending as a line with points
  geom_line(aes(y = (mmc_spending / 1e9) * (max_enrollee_mmc / max_spending_mmc),
                group = 1, color = "Spending")) +
  geom_point(aes(y = (mmc_spending / 1e9) * (max_enrollee_mmc / max_spending_mmc)), color = "black") +
  # Left y-axis for enrollee_months
  scale_y_continuous(
    name = "Enrollee Months (millions)\n",
    labels = scales::label_comma(),
    breaks = breaks_pretty(n = 8),
    sec.axis = sec_axis(~ . * (max_spending_mmc / max_enrollee_mmc), 
                        name = "MMC Spending\n", 
                        breaks = breaks_pretty(n = 8),
                        labels = scales::label_dollar(scale = 1, suffix = "B"))) +
  scale_fill_manual(values = "#B9CDE5") +
  scale_color_manual(values = "black") +
  # Formatting
  labs(x = "\nFiscal Year",
       title = "Medicaid Managed Care Enrollment and Spending, by Fiscal Year",
       subtitle = paste0("New York State (",
                         # mmc_global_cap_analysis_data_v1 %>% slice_min(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                         # " to ",
                         # mmc_global_cap_analysis_data_v1 %>% slice_max(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                         "Since April 2011",
                         ")"),
       color = "",
       fill = "",
       caption = paste0("Data Sources: New York State Department of Health Medicaid Managed Care Enrollment Reports, and Medicaid Global Spending Cap Updates\nData as of: ",format(Sys.Date(),"%B %d, %Y"))) +
  theme_light(base_size = 16) +
  isaac_step2_graph_theme +
  theme(legend.text = element_text(face = "bold"))

mmc_enrollment_vs_spending_table_data %>%
  select(fiscal_yr,
         n_months,
         enrollee_months,
         mean_number_of_enrollees,
         end_of_period_enrollment,
         enrollee_month_delta,
         enrollee_month_delta_pct,
         mmc_spending,
         mmc_spending_delta,
         mmc_spending_delta_pct,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta,
         mean_spend_per_enrollee_delta_pct) %>% 
  mutate(fiscal_yr = ifelse(n_months != 12,paste0(fiscal_yr," (YTD)"),fiscal_yr),
         across(.cols = -c(fiscal_yr,
                           n_months,
                           enrollee_months,
                           mean_number_of_enrollees,
                           end_of_period_enrollment,
                           enrollee_month_delta,
                           enrollee_month_delta_pct),
                .fns = ~ ifelse(n_months != 12,NA,.x))) %>% 
  rename("State Fiscal Year" = "fiscal_yr",
         "Months Reported" = "n_months",
         "Enrollee Months" = "enrollee_months",
         "Average Annual Enrollment" = "mean_number_of_enrollees",
         "End of Period Enrollment" = "end_of_period_enrollment",
         "Enrollee Months Annual Change" = "enrollee_month_delta",
         "Enrollee Months % Growth" = "enrollee_month_delta_pct",
         "State Spending (Millions)" = "mmc_spending",
         "State Spending Annual Change (Millions)" = "mmc_spending_delta",
         "State Spending Annual Change %" = "mmc_spending_delta_pct",
         "Average Annual Spending per Enrollee" = "mean_spend_per_enrollee",
         "Average Annual Spending per Enrollee Annual Change" = "mean_spend_per_enrollee_delta",
         "Average Annual Per Capita Spending % Growth" = "mean_spend_per_enrollee_delta_pct") %>% 
  write_csv("mmc_table_data.csv")

mmc_enrollment_vs_spending_table_data %>%
  filter(n_months == 12) %>% 
  mutate(data_type = "Actual",
         fiscal_yr_n = fiscal_yr %>% str_remove("FY") %>% as.numeric(),
         fiscal_yr = ifelse(n_months != 12,paste0(fiscal_yr," (YTD)"),fiscal_yr)) %>% 
  arrange(fiscal_yr) %>% 
  mutate(mean_number_of_enrollees_delta_pct = (mean_number_of_enrollees - lag(mean_number_of_enrollees)) / lag(mean_number_of_enrollees)) %>% 
  select(data_type,
         fiscal_yr,
         mmc_spending,
         mean_number_of_enrollees,
         mean_number_of_enrollees_delta_pct,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta_pct) %>% 
  rename("Data Type" = "data_type",
         "Fiscal Year" = "fiscal_yr",
         "State Spending" = "mmc_spending",
         "Average Annual Enrollment" = "mean_number_of_enrollees",
         "Average Annual Enrollment Change %" = "mean_number_of_enrollees_delta_pct",
         "Average Annual Spending per Enrollee" = "mean_spend_per_enrollee",
         "Average Annual Spending per Enrollee Annual Change %" = "mean_spend_per_enrollee_delta_pct") %>% 
  write_csv("mmc_projection_data.csv")
```

<br>

```{r mmc_enrollment_vs_spending_table}
mmc_enrollment_vs_spending_table %>% 
  tab_source_note(
    # mmc_enrollment_vs_spending_table_data %>%
    #   mutate(across(.cols = ends_with("pct"),
    #                 .fns = ~ percent(.x,accuracy = 0.1)),
    #          across(.cols = contains("spend") & !ends_with("pct"),
    #                 .fns = ~ dollar(.x,accuracy = 1)),
    #          across(.cols = c(enrollee_months,enrollee_month_delta,mean_number_of_enrollees),
    #                 .fns = ~ comma(.x,accuracy = 1))) %>% 
    #   rename("Fiscal Year" = "fiscal_yr",
    #          "Months Reported" = "n_months",
    #          "Enrollee Months" = "enrollee_months",
    #          "Average Monthly Change %" = "mean_monthly_change_pct",
    #          "Enrollee Months Annual Change" = "enrollee_month_delta",
    #          "Enrollee Months Annual Change %" = "enrollee_month_delta_pct",
    #          "State Spending" = "mmc_spending",
    #          "State Spending Annual Change" = "mmc_spending_delta",
    #          "State Spending Annual Change %" = "mmc_spending_delta_pct",
    #          "Average Monthly Number of Enrollees" = "mean_number_of_enrollees",
    #          "Average Annual Spending per Enrollee" = "mean_spend_per_enrollee",
    #          "Average Annual Spending per Enrollee Annual Change" = "mean_spend_per_enrollee_delta",
    #          "Average Annual Spending per Enrollee Annual Change %" = "mean_spend_per_enrollee_delta_pct") %>% 
    #   download_this(
    #     output_name = paste0("mmc_enrollment_vs_spending__",str_replace_all(Sys.Date(),"-","_")),
    #     output_extension = ".xlsx",
    #     file_path = "https://isaacmichaels.github.io/s2p/mmc_enrollment_vs_spending__2025_01_10.xlsx",
    #     button_label = "Download Excel",
    #     button_type = "default",
    #   )
    download_link(
      link = "https://github.com/steptwopolicy/s2p_data_analyses/raw/refs/heads/main/mmc_enrollment_vs_spending.xlsx",
      button_label = "Download Interactive Excel Workbook",
      button_type = "default",
      class = "button_download"
    )
  )
```

The interactive Excel workbook allows users to enter growth-rate predictions, and model projections.  
**Data Sources:** New York State Department of Health [Medicaid Managed Care Enrollment Reports](https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/), and [Medicaid Global Spending Cap Updates
](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/)

<br>

```{r mmc_enrollment_vs_spending_graph}
mmc_enrollment_vs_spending_graph
```

**Data Sources:** New York State Department of Health [Medicaid Managed Care Enrollment Reports](https://www.health.ny.gov/health_care/managed_care/reports/enrollment/monthly/), and [Medicaid Global Spending Cap Updates
](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/)

<br>
