---
output:
  html_document:
    toc: false
params:
  project_directory: "/Users/isaac/Dropbox/S2P Datasets Shared Folder/MLTC_MMC_NH_data/"
  use_local_data: FALSE
---
<style>
.button_download {
font-size: 16px;
font-family: "Arial";
color: black;
background-color: #C7C9D8;
}
</style>

```{r setup_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, include = TRUE, echo = FALSE, message = FALSE, warning = FALSE, error = TRUE, fig.width = 12, fig.height = 8, cache = FALSE)
```

```{r setup_libraries, include=FALSE}
library(tidyverse)
library(furrr)
library(scales)
library(gt)
library(pdftools)
library(httr)
library(downloadthis)
data_dir <- params$project_directory
```

```{r isaac_step2_graph_theme, include=FALSE}
isaac_step2_graph_theme <- theme(plot.title.position = "plot",
                                 plot.title = element_text(face = "bold"),
                                 legend.position = "top",
                                 strip.text = element_text(color = "black"),
                                 legend.title = element_text(face = "bold",hjust = 0.5),
                                 axis.title.y = element_text(face = "bold",angle = 90),
                                 axis.title.x = element_text(face = "bold"))
```

```{r setup_medicaid_enrollment_data_ingestion, include=FALSE}
if(params$use_local_data == FALSE){
  # https://health.data.ny.gov/Health/Medicaid-Program-Enrollment-by-Month-Beginning-200/m4hz-kzn3/about_data
  nys_medicaid <- read_csv("https://health.data.ny.gov/api/views/m4hz-kzn3/rows.csv?accessType=DOWNLOAD") %>% janitor::clean_names()
  
  write_csv(nys_medicaid,paste0(data_dir,'mmc_mltc/data/Medicaid-Program-Enrollment-by-Month-Beginning-2009.csv'))
} else{
  nys_medicaid <- read_csv(paste0(data_dir,'mmc_mltc/data/Medicaid-Program-Enrollment-by-Month-Beginning-2009.csv'))
}
```

```{r setup_medicaid_spending_data_ingestion, include=FALSE}
if(params$use_local_data == FALSE){
  # https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/
  try_downloading_report <- function(){
    # Use GET with a User-Agent header
    response <- GET(pdf_url, user_agent("Mozilla/5.0 (Windows NT 10.0; Win64; x64) AppleWebKit/537.36 (KHTML, like Gecko) Chrome/121.0.0.0 Safari/537.36"))
    
    # Check response status
    if (status_code(response) == 200) {
      writeBin(content(response, "raw"), pdf_file)
      message("File downloaded successfully.")
    } else {
      message("Failed to download file. HTTP status: ", status_code(response))
    }
  }
  
  # Download Global Cap Quarterly Reports
  for(i in c(2011:(as.numeric(format(Sys.Date(),"%Y"))))){
    for(j in c(1:4)){
      qrt_number <- case_when(j == 1 ~ "1st",
                              j == 2 ~ "2nd",
                              j == 3 ~ "3rd",
                              j == 4 ~ "4th")
      
      pdf_url <- paste0("https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/monthly/sfy_",i,"-",i+1,"/docs/",qrt_number,"_qtr_rpt.pdf")
      
      pdf_file <- paste0(data_dir,"global_cap_reports/nys_medicaid_",i,"-",i+1,"_q",j,"report.pdf")
      
      if(file.exists(pdf_file) == FALSE){
        try_downloading_report()
      } else{
        message("File already downloaded")
      }
    }
  }
  
  # Download Global Cap Monthly Reports
  for(i in c(2011:(as.numeric(format(Sys.Date(),"%Y"))))){
    for(j in c(month.name,month.abb,"Sept")){
      for(k in c(0,1)){
        pdf_url <- paste0("https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/monthly/sfy_",i,"-",i+1,"/docs/",j,"_",i+k,"_report.pdf") %>% 
          str_to_lower()
        
        pdf_file <- paste0(data_dir,"global_cap_reports/nys_medicaid_",i,"-",i+1,"_month_",j,"_",i+k,"_report.pdf")
        
        if(file.exists(pdf_file) == FALSE){
          try_downloading_report()
        } else{
          message("File already downloaded")
        }
      }
    }
  }
  
  report_names <- list.files(paste0(data_dir,"global_cap_reports/"),
                             pattern = ".pdf",
                             full.names = FALSE,
                             recursive = FALSE) %>% 
    # Unusually formatted non-Q4 report
    setdiff("nys_medicaid_2024-2025_q2report.pdf")
  
  report_paths <- list.files(paste0(data_dir,"global_cap_reports/"),
                             pattern = ".pdf",
                             full.names = TRUE,
                             recursive = FALSE) %>% 
    # Unusually formatted non-Q4 report
    setdiff("/Users/isaac/Dropbox/S2P Datasets Shared Folder/MLTC_MMC_NH_data/global_cap_reports//nys_medicaid_2024-2025_q2report.pdf")
  
  for(i in c(1:length(report_names))){
    report_name <- report_names[i]
    
    # Extract text from the PDF
    pdf_text <- pdf_text(report_paths[i])
    
    report_table_page <- c(1:length(pdf_text))[pdf_text %>% str_to_lower() %>% str_detect("actual") &
                                                 pdf_text %>% str_to_lower() %>% str_detect("variance") &
                                                 (pdf_text %>% str_to_lower() %>% str_detect("category of service") | pdf_text %>% str_to_lower() %>% str_detect("category of spending") | pdf_text %>% str_to_lower() %>% str_detect("quarter 2")) &
                                                 pdf_text %>% str_detect("TOTAL") &
                                                 pdf_text %>% str_to_lower() %>% str_detect("table of contents",negate = TRUE)] %>% 
      min()
    
    pdf_table <- pdf_text[report_table_page]
    
    dollar_multiplier <- case_when(pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("in millions") ~ 1000000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("in thousands") ~ 1000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("millions") ~ 1000000,
                                   pdf_table %>% as.character() %>% str_to_lower() %>% str_detect("thousands") ~ 1000)
    
    # Parse the table from the extracted text
    table_data <- str_split(pdf_table,"\n")[[1]] %>% 
      # Clean and structure the data
      str_trim() %>%
      as_tibble()
    
    table_start_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_to_lower() %>% str_sub(1,19) == "category of service" |
               value %>% str_to_lower() %>% str_sub(1,20) == "category of spending") %>% 
      dplyr::pull(rowid)
    
    first_data_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_count("\\$") == 3,
             rowid > table_start_row) %>% 
      dplyr::pull(rowid) %>%
      min()
    
    table_end_row <- table_data %>% 
      rowid_to_column() %>% 
      filter(value %>% str_sub(1,5) == "TOTAL") %>% 
      dplyr::pull(rowid)
    
    header_rows <- table_data %>% 
      slice(table_start_row:first_data_row - 1) %>% 
      as.character() %>% 
      paste(collapse = " ") %>% 
      str_to_lower()
    
    table_col_names <- case_when(str_detect(header_rows,"target") ~ c("category_of_service","target","actual","variance"),
                                 str_detect(header_rows,"estimated") ~ c("category_of_service","estimated","actual","variance"))
    
    table_clean <- table_data %>%
      slice(first_data_row:table_end_row) %>% 
      separate(col = value, into = c("Column1", "Column2", "Column3", "Column4"), sep = "\\s{2,}", fill = "right")
    
    names(table_clean) <- table_col_names
    
    table_clean <- table_clean %>% 
      mutate(report = report_name %>% str_remove(".pdf"),
             multiplier = dollar_multiplier)
    
    if(i == 1){
      nys_medicaid_spending <- table_clean
    } else{
      nys_medicaid_spending <- bind_rows(nys_medicaid_spending,table_clean)
    }
    paste0(i," of ",length(report_names),"(",scales::percent(i/length(report_names),accuracy = 0.1),") | ",report_name) %>% print()
  }
  write_csv(nys_medicaid_spending,paste0(data_dir,'mmc_mltc/data/Medicaid-Spending-by-Fiscal_Year.csv'))
} else{
  nys_medicaid_spending <- read_csv(paste0(data_dir,'mmc_mltc/data/Medicaid-Spending-by-Fiscal_Year.csv'))
}
```

```{r setup_medicaid_enrollment_vs_spending_analysis, include=FALSE}
mltc_analysis <- nys_medicaid %>% 
  filter(managed_care_vs_fee_for_service == "MMC",
         plan_type %in% c("PACE",
                          "PARTIAL MLTC",
                          "MAP")) %>% 
  mutate(month_year = paste(eligibility_year,eligibility_month,"01",sep = "-") %>% as.Date(),
         year = paste(eligibility_year,"01","01",sep = "-") %>% as.Date())

mmc_global_cap_analysis <- nys_medicaid_spending %>% 
  mutate(reference = ifelse(is.na(target),estimated,target),
         ref_type = ifelse(is.na(target),"estimated","target"),
         across(.cols = c("reference","actual"),
                .fns = ~ (.x %>%
                            # Remove $ and ,
                            str_remove_all("[\\$,]") %>%
                            # Replace parentheses with a negative sign
                            str_replace_all("\\((.*)\\)", "-\\1") %>%
                            # Convert to numeric
                            as.numeric()) * multiplier),
         difference = actual - reference) %>% 
  select(category_of_service,
         ref_type,
         reference,
         actual,
         report) %>% 
  # Only include year-end reports in the analysis
  filter(str_detect(report,"q4|Mar")) %>% 
  mutate(fiscal_yr = str_sub(report,14,22),
         type = case_when(category_of_service %in% c("Long Term Managed Care") ~ "Long Term Care")) %>% 
  select(-report)

mmc_global_cap_analysis_data_v2 <- mmc_global_cap_analysis %>% 
  mutate(fiscal_yr_dt = fiscal_yr %>% str_sub(1,4) %>% paste("01","01",sep="-") %>% as.Date()) %>% 
  select(-ref_type,
         -reference) %>% 
  filter(!is.na(type)) %>% 
  group_by(fiscal_yr,
           fiscal_yr_dt,
           type) %>%
  summarise(actual = sum(actual)) %>% 
  ungroup()

mltc_enrollment_vs_spending_table_data <- mltc_analysis %>% 
  mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                               TRUE ~ eligibility_year)) %>% 
  group_by(fiscal_yr,plan_type) %>% 
  summarise(enrollee_months = sum(number_of_recipients)) %>% 
  ungroup() %>% 
  pivot_wider(names_from = plan_type,
              values_from = enrollee_months) %>% 
  left_join(mltc_analysis %>% 
              mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                                           TRUE ~ eligibility_year)) %>% 
              group_by(fiscal_yr,eligibility_month) %>% 
              summarise(enrollee_months = sum(number_of_recipients)) %>% 
              ungroup() %>% 
              group_by(fiscal_yr) %>% 
              slice_max(eligibility_month) %>% 
              ungroup() %>% 
              select(fiscal_yr,enrollee_months) %>% 
              rename("end_of_period_enrollment" = "enrollee_months"),
            by = "fiscal_yr") %>% 
  left_join(mmc_global_cap_analysis_data_v2 %>% 
              filter(type == "Long Term Care") %>% 
              select(fiscal_yr,
                     actual) %>% 
              mutate(fiscal_yr = fiscal_yr %>% str_sub(1,4) %>% as.numeric()) %>% 
              rename("mltc_spending" = "actual"),
            by = "fiscal_yr") %>% 
  left_join(mltc_analysis %>%
              mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                                           TRUE ~ eligibility_year)) %>%
              group_by(fiscal_yr,eligibility_month) %>%
              summarise(enrollee_months = sum(number_of_recipients)) %>%
              ungroup() %>%
              group_by(fiscal_yr) %>%
              summarise(n_months = n(),
                        mean_number_of_enrollees = mean(enrollee_months)) %>%
              ungroup(),
            by = "fiscal_yr") %>% 
  left_join(mltc_analysis %>% 
              mutate(fiscal_yr = case_when(eligibility_month <= 3 ~ eligibility_year - 1,
                                           TRUE ~ eligibility_year)) %>% 
              group_by(fiscal_yr,eligibility_month) %>% 
              summarise(enrollee_months = sum(number_of_recipients)) %>% 
              ungroup() %>% 
              arrange(desc(fiscal_yr),
                      desc(eligibility_month)) %>% 
              mutate(enrollee_months_delta = enrollee_months  - lead(enrollee_months),
                     enrollee_months_delta_pct = enrollee_months_delta / lead(enrollee_months,n=1)) %>% 
              group_by(fiscal_yr) %>% 
              summarise(mean_monthly_change_pct = mean(enrollee_months_delta_pct)) %>% 
              ungroup(),
            by = "fiscal_yr") %>% 
  arrange(desc(fiscal_yr)) %>% 
  filter(fiscal_yr >= 2014) %>% 
  mutate(fiscal_yr_dt = fiscal_yr %>% str_sub(1,4) %>% paste("01","01",sep="-") %>% as.Date(),
         fiscal_yr = fiscal_yr_dt %>% format("%Y") %>% as.numeric(),
         fiscal_yr = paste(fiscal_yr,str_sub(fiscal_yr + 1,3,4),sep = "-"),
         fiscal_yr = paste0("FY",str_sub(fiscal_yr,-2)),
         enrollee_months = MAP + PACE + `PARTIAL MLTC`,
         mltc_spending_delta = mltc_spending  - lead(mltc_spending),
         mltc_spending_delta_pct = mltc_spending_delta / lead(mltc_spending,n=1),,
         spending_per_enrollee_month = mltc_spending / enrollee_months,
         spending_per_enrollee_month_delta = spending_per_enrollee_month  - lead(spending_per_enrollee_month),
         spending_per_enrollee_month_delta_pct = spending_per_enrollee_month_delta / lead(spending_per_enrollee_month,n=1),
         enrollee_month_delta = enrollee_months - lead(enrollee_months),
         enrollee_month_delta_pct = enrollee_month_delta / lead(enrollee_months,n=1),
         mean_spend_per_enrollee = mltc_spending / mean_number_of_enrollees,
         mean_spend_per_enrollee_delta = mean_spend_per_enrollee - lead(mean_spend_per_enrollee),
         mean_spend_per_enrollee_delta_pct = mean_spend_per_enrollee_delta / lead(mean_spend_per_enrollee,n=1)) %>% 
  select(fiscal_yr,
         n_months,
         MAP,
         PACE,
         `PARTIAL MLTC`,
         enrollee_months,
         end_of_period_enrollment,
         mean_monthly_change_pct,
         enrollee_month_delta,
         enrollee_month_delta_pct,
         mltc_spending,
         mltc_spending_delta,
         mltc_spending_delta_pct,
         mean_number_of_enrollees,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta,
         mean_spend_per_enrollee_delta_pct)

mltc_enrollment_vs_spending_table <- mltc_enrollment_vs_spending_table_data %>% 
  gt() %>% 
  tab_header(title = md("**Managed Long Term Care Enrollment and Spending, by Fiscal Year**"),
             subtitle = paste0("New York State (",
                               mmc_global_cap_analysis_data_v2 %>% filter(type == "Long Term Care") %>% slice_min(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                               " to ",
                               mmc_global_cap_analysis_data_v2 %>% filter(type == "Long Term Care") %>% slice_max(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                               ")")) %>%
  fmt_integer(c(n_months,
                MAP,
                PACE,
                `PARTIAL MLTC`,
                enrollee_months,
                enrollee_month_delta,
                mean_number_of_enrollees,
                end_of_period_enrollment)) %>% 
  fmt_currency(c(mltc_spending,
                 mltc_spending_delta,
                 mean_spend_per_enrollee,
                 mean_spend_per_enrollee_delta),
               decimals = 0) %>% 
  fmt_percent(c(enrollee_month_delta_pct,
                mltc_spending_delta_pct,
                mean_spend_per_enrollee_delta_pct,
                mean_monthly_change_pct),
              decimals = 1) %>% 
  tab_spanner(label = "Enrollee Months, by Plan Type",
              columns = c(MAP,
                          PACE,
                          `PARTIAL MLTC`)) %>% 
  tab_spanner(label = "Total Enrollment",
              columns = c(enrollee_months,
                          mean_monthly_change_pct,
                          enrollee_month_delta,
                          enrollee_month_delta_pct,
                          end_of_period_enrollment)) %>% 
  tab_spanner(label = "State Spending",
              columns = c(mltc_spending,
                          mltc_spending_delta,
                          mltc_spending_delta_pct)) %>% 
  tab_spanner(label = "Average Annual Spending per Enrollee",
              columns = c(mean_number_of_enrollees,
                          mean_spend_per_enrollee,
                          mean_spend_per_enrollee_delta,
                          mean_spend_per_enrollee_delta_pct)) %>% 
  cols_label(fiscal_yr = "Fiscal Year",
             n_months = "Months Reported",
             `PARTIAL MLTC` = "MLTC",
             enrollee_months = "Enrollee Months",
             mean_monthly_change_pct = "Average Monthly Change %",
             enrollee_month_delta = "Annual Change",
             enrollee_month_delta_pct = "Annual Change %",
             end_of_period_enrollment = "End of Period Enrollment",
             mltc_spending = "Spending",
             mltc_spending_delta = "Annual Change",
             mltc_spending_delta_pct = "Annual Change %",
             mean_number_of_enrollees = "Average Monthly Number of Enrollees",
             mean_spend_per_enrollee = "Spending",
             mean_spend_per_enrollee_delta = "Annual Change",
             mean_spend_per_enrollee_delta_pct = "Annual Change %") %>% 
  cols_align(align = "left", columns = fiscal_yr) %>% 
  tab_style(style = cell_text(weight = "bold"),
            locations = list(cells_column_labels(),
                             cells_column_spanners())) %>% 
  tab_source_note(source_note = md("**Data Sources:** Health Data NY; NYSDOH Global Spending Cap
Reports")) %>% 
  tab_source_note(source_note = md(paste0("**Data as of:** ",format(Sys.Date(),"%B %d, %Y")))) %>% 
  tab_options(table.font.size = px(10),
              table.font.names = "Helvetica",
              data_row.padding = px(2),
              table.width = pct(95),
              container.width = px(1000))

# Calculate scaling factors for normalization
max_enrollee_mltc <- max(mltc_enrollment_vs_spending_table_data$enrollee_months,na.rm=TRUE) / 1e6
max_spending_mltc <- max(mltc_enrollment_vs_spending_table_data$mltc_spending,na.rm=TRUE) / 1e9

mltc_enrollment_vs_spending_graph <- mltc_enrollment_vs_spending_table_data %>% 
  select(fiscal_yr,
         enrollee_months,
         mltc_spending) %>% 
  # Create the plot
  ggplot(aes(x = fiscal_yr)) +
  # Plot enrollee_months as bars
  geom_bar(aes(y = enrollee_months / 1e6, fill = "Enrollment"),
           stat = "identity", alpha = 0.9) +
  # Plot mltc_spending as a line with points
  geom_line(aes(y = (mltc_spending / 1e9) * (max_enrollee_mltc / max_spending_mltc),
                group = 1, color = "Spending")) +
  geom_point(aes(y = (mltc_spending / 1e9) * (max_enrollee_mltc / max_spending_mltc)), color = "black") +
  # Left y-axis for enrollee_months
  scale_y_continuous(
    name = "Enrollee Months (millions)\n",
    labels = scales::label_comma(),
    breaks = breaks_pretty(n = 8),
    sec.axis = sec_axis(~ . * (max_spending_mltc / max_enrollee_mltc), 
                        name = "MLTC Spending\n", 
                        breaks = breaks_pretty(n = 8),
                        labels = scales::label_dollar(scale = 1, suffix = "B"))) +
  scale_fill_manual(values = "#B9CDE5") +
  scale_color_manual(values = "black") +
  # Formatting
  labs(x = "\nFiscal Year",
       title = "Medicaid Managed Long Term Care Enrollment and Spending, by Fiscal Year",
       subtitle = paste0("New York State (",
                         mmc_global_cap_analysis_data_v2 %>% filter(type == "Long Term Care") %>% slice_min(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                         " to ",
                         mmc_global_cap_analysis_data_v2 %>% filter(type == "Long Term Care") %>% slice_max(fiscal_yr_dt) %>% dplyr::pull(fiscal_yr) %>% unique(),
                         ")"),
       color = "",
       fill = "",
       caption = paste0("Data Sources: Health Data NY; NYSDOH Global Spending Cap Reports\nData as of: ",format(Sys.Date(),"%B %d, %Y"))) +
  theme_light(base_size = 16) +
  isaac_step2_graph_theme +
  theme(legend.text = element_text(face = "bold"))

mltc_trends_by_plan_name <- mltc_analysis %>% 
  group_by(plan_name,month_year,plan_type) %>% 
  summarise(enrolled = sum(number_of_recipients)) %>% 
  ungroup() %>% 
  # mutate(plan_name = plan_name %>% factor(levels = mltc_analysis %>% 
  #                                           group_by(plan_name) %>% 
  #                                           summarise(enrolled = sum(number_of_recipients)) %>% 
  #                                           ungroup() %>% 
  #                                           arrange(desc(enrolled)) %>% 
  #                                           dplyr::pull(plan_name))) %>% 
  ggplot(aes(x = month_year,
             y = enrolled,
             color = plan_type)) + 
  geom_line(linewidth = 1.25) +
  labs(title = "Medicaid Managed Long Term Care Plans, Statewide Enrollment by Plan Name, Plan Type, and Month",
       subtitle = paste0("New York State (",
                         format(min(mltc_analysis$month_year),"%b %Y"),
                         " to ",
                         format(max(mltc_analysis$month_year),"%b %Y"),
                         ")"),
       x = "",
       y = "Number of Enrollees\n",
       color = "Plan Type",
       caption = paste0("Data Source: Health Data NY\nData as of: ",format(Sys.Date(),"%B %d, %Y"))) +
  scale_x_date(breaks = breaks_pretty(n = 4)) +
  scale_y_continuous(labels = comma_format(accuracy = 1),
                     limits = c(0,NA)) +
  scale_color_manual(values = c("#1b9e77","#d95f02","#7570b3")) +
  theme_light(base_size = 12) +
  isaac_step2_graph_theme +
  facet_wrap(~plan_name,
             labeller = label_wrap_gen(width = 20),
             ncol = 5,
             axes = "all",
             scales = "free_y")

mltc_trends_by_plan_name_table <- mltc_analysis %>% 
  group_by(plan_name,year,plan_type) %>% 
  summarise(enrolled = sum(number_of_recipients)) %>% 
  ungroup() %>% 
  mutate(year = year %>% format("%Y") %>% as.numeric()) %>% 
  pivot_wider(names_from = year,
              values_from = enrolled,
              names_sort = TRUE,
              values_fill = 0) %>% 
  relocate(plan_type,plan_name) %>% 
  arrange(plan_type,plan_name) %>% 
  gt(groupname_col = "plan_type",
     row_group_as_column = TRUE) %>% 
  tab_header(title = md("**Medicaid Managed Long Term Care Plan<br>Member Months, by Plan Type, Plan Name, and Year**"),
             subtitle = paste0("New York State (",
                               format(min(mltc_analysis$month_year),"%b %Y"),
                               " to ",
                               format(max(mltc_analysis$month_year),"%b %Y"),
                               ")")) %>% 
  opt_row_striping() %>% 
  fmt_integer(is.numeric) %>% 
  cols_label(plan_name = "Plan Name",
             plan_type = "Plan Type") %>% 
  cols_align(align = "left", columns = plan_name) %>% 
  tab_style(style = cell_text(weight = "bold"),
            locations = cells_column_labels()) %>% 
  tab_source_note(source_note = md("Scroll right to see additional years")) %>% 
  tab_source_note(source_note = md("**Data Source:** Health Data NY")) %>% 
  tab_source_note(source_note = md(paste0("**Data as of:** ",format(Sys.Date(),"%B %d, %Y")))) %>% 
  tab_options(table.font.size = px(10),
              table.font.names = "Helvetica",
              data_row.padding = px(2))

mltc_enrollment_vs_spending_table_data %>%
  select(fiscal_yr,
         n_months,
         MAP,
         PACE,
         `PARTIAL MLTC`,
         enrollee_months,
         mean_number_of_enrollees,
         end_of_period_enrollment,
         enrollee_month_delta,
         enrollee_month_delta_pct,
         mltc_spending,
         mltc_spending_delta,
         mltc_spending_delta_pct,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta,
         mean_spend_per_enrollee_delta_pct) %>% 
  mutate(fiscal_yr = ifelse(n_months != 12,paste0(fiscal_yr," (YTD)"),fiscal_yr),
         across(.cols = -c(fiscal_yr,
                           n_months,
                           MAP,
                           PACE,
                           `PARTIAL MLTC`,
                           enrollee_months,
                           mean_number_of_enrollees,
                           end_of_period_enrollment),
                .fns = ~ ifelse(n_months != 12,NA,.x))) %>% 
  rename("State Fiscal Year" = "fiscal_yr",
         "Months Reported" = "n_months",
         "MAP Enrollee Months" = "MAP",
         "PACE Enrollee Months" = "PACE",
         "PARTIAL MLTC Enrollee Months" = "PARTIAL MLTC",
         "Total Enrollee Months" = "enrollee_months",
         "Average Annual Enrollment" = "mean_number_of_enrollees",
         "End of Period Enrollment" = "end_of_period_enrollment",
         "Enrollee Months Annual Change" = "enrollee_month_delta",
         "Enrollee Months % Growth" = "enrollee_month_delta_pct",
         "State Spending (Millions)" = "mltc_spending",
         "State Spending Annual Change (Millions)" = "mltc_spending_delta",
         "State Spending Annual Change %" = "mltc_spending_delta_pct",
         "Average Annual Spending per Enrollee" = "mean_spend_per_enrollee",
         "Average Annual Spending per Enrollee Annual Change" = "mean_spend_per_enrollee_delta",
         "Average Annual Per Capita Spending % Growth" = "mean_spend_per_enrollee_delta_pct") %>% 
  write_csv("mltc_table_data.csv")

mltc_enrollment_vs_spending_table_data %>%
  filter(n_months == 12) %>% 
  mutate(data_type = "Actual",
         fiscal_yr_n = fiscal_yr %>% str_remove("FY") %>% as.numeric(),
         fiscal_yr = ifelse(n_months != 12,paste0(fiscal_yr," (YTD)"),fiscal_yr)) %>% 
  arrange(fiscal_yr) %>% 
  mutate(mean_number_of_enrollees_delta_pct = (mean_number_of_enrollees - lag(mean_number_of_enrollees)) / lag(mean_number_of_enrollees)) %>% 
  select(data_type,
         fiscal_yr,
         mltc_spending,
         mean_number_of_enrollees,
         mean_number_of_enrollees_delta_pct,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta_pct) %>% 
  rename("Data Type" = "data_type",
         "Fiscal Year" = "fiscal_yr",
         "State Spending" = "mltc_spending",
         "Average Annual Enrollment" = "mean_number_of_enrollees",
         "Average Annual Enrollment Change %" = "mean_number_of_enrollees_delta_pct",
         "Average Annual Spending per Enrollee" = "mean_spend_per_enrollee",
         "Average Annual Spending per Enrollee Annual Change %" = "mean_spend_per_enrollee_delta_pct") %>% 
  write_csv("mltc_projection_data.csv")
```

```{r setup_excel_workbook_data, include=FALSE}
datatable_page <- mltc_enrollment_vs_spending_table_data %>% 
  select(fiscal_yr,
         n_months,
         MAP,
         PACE,
         `PARTIAL MLTC`,
         enrollee_months,
         mean_number_of_enrollees,
         end_of_period_enrollment,
         enrollee_month_delta,
         enrollee_month_delta_pct,
         mltc_spending,
         mltc_spending_delta,
         mltc_spending_delta_pct,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta,
         mean_spend_per_enrollee_delta_pct)

projection_data_page <- mltc_enrollment_vs_spending_table_data %>% 
  select(fiscal_yr,
         mltc_spending,
         mean_number_of_enrollees,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta_pct) %>% 
  arrange(desc(fiscal_yr)) %>% 
  mutate(mean_number_of_enrollees_delta = mean_number_of_enrollees - lead(mean_number_of_enrollees),
         mean_number_of_enrollees_delta_pct = mean_number_of_enrollees_delta / lead(mean_number_of_enrollees,n=1)) %>% 
  select(fiscal_yr,
         mltc_spending,
         mean_number_of_enrollees,
         mean_number_of_enrollees_delta_pct,
         mean_spend_per_enrollee,
         mean_spend_per_enrollee_delta_pct) %>% 
  arrange(fiscal_yr)
```

<br>

### Enrollment and Spending

```{r mltc_enrollment_vs_spending_table}
mltc_enrollment_vs_spending_table %>% 
  tab_source_note(
    # mltc_enrollment_vs_spending_table_data %>%
    #   mutate(across(.cols = ends_with("pct"),
    #                 .fns = ~ percent(.x,accuracy = 0.1)),
    #          across(.cols = contains("spend") & !ends_with("pct"),
    #                 .fns = ~ dollar(.x,accuracy = 1)),
    #          across(.cols = c(MAP,PACE,`PARTIAL MLTC`,
    #                           enrollee_months,enrollee_month_delta,mean_number_of_enrollees),
    #                 .fns = ~ comma(.x,accuracy = 1))) %>% 
    #   rename("Fiscal Year" = "fiscal_yr",
    #          "Months Reported" = "n_months",
    #          "Enrollee Months" = "enrollee_months",
    #          "Average Monthly Change %" = "mean_monthly_change_pct",
    #          "Enrollee Months Annual Change" = "enrollee_month_delta",
    #          "Enrollee Months Annual Change %" = "enrollee_month_delta_pct",
    #          "State Spending" = "mltc_spending",
    #          "State Spending Annual Change" = "mltc_spending_delta",
    #          "State Spending Annual Change %" = "mltc_spending_delta_pct",
    #          "Average Monthly Number of Enrollees" = "mean_number_of_enrollees",
    #          "Average Annual Spending per Enrollee" = "mean_spend_per_enrollee",
    #          "Average Annual Spending per Enrollee Annual Change" = "mean_spend_per_enrollee_delta",
    #          "Average Annual Spending per Enrollee Annual Change %" = "mean_spend_per_enrollee_delta_pct") %>% 
    #   download_this(
    #     output_name = paste0("mltc_enrollment_vs_spending__",str_replace_all(Sys.Date(),"-","_")),
    #     output_extension = ".xlsx",
    #     button_label = "Download Excel",
    #     button_type = "default",
    #   )
    download_link(
      link = "https://github.com/steptwopolicy/s2p_data_analyses/raw/refs/heads/main/mltc_enrollment_vs_spending.xlsx",
      button_label = "Download Interactive Excel Workbook",
      button_type = "default",
      class = "button_download"
    )
  )
```

The interactive Excel workbook allows users to enter growth-rate predictions, and model projections.  
**Data Sources:** [Health Data NY](https://health.data.ny.gov/Health/Medicaid-Program-Enrollment-by-Month-Beginning-200/m4hz-kzn3/about_data); [New York State Department of Health](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/)

<br>

```{r mltc_enrollment_vs_spending_graph}
mltc_enrollment_vs_spending_graph
```
**Data Sources:** [Health Data NY](https://health.data.ny.gov/Health/Medicaid-Program-Enrollment-by-Month-Beginning-200/m4hz-kzn3/about_data); [New York State Department of Health](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/)

<br>

### Enrollment, by Plan Name

```{r mltc_trends_by_plan_name_table}
mltc_trends_by_plan_name_table
```
**Data Sources:** [Health Data NY](https://health.data.ny.gov/Health/Medicaid-Program-Enrollment-by-Month-Beginning-200/m4hz-kzn3/about_data); [New York State Department of Health](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/)

<br>

```{r mltc_trends_by_plan_name,fig.height=20}
mltc_trends_by_plan_name
```
**Data Sources:** [Health Data NY](https://health.data.ny.gov/Health/Medicaid-Program-Enrollment-by-Month-Beginning-200/m4hz-kzn3/about_data); [New York State Department of Health](https://www.health.ny.gov/health_care/medicaid/regulations/global_cap/)

<br>

***

### Technical Notes

1. <i>MLTC</i> = Partial cap managed long term care.
2. Spending data in this analysis are based on the amounts reported under the 'Long Term Managed Care' category in the New York State Medicaid Global Cap Reports. This category reflects state-reported expenditures specific to Medicaid Managed Long-Term Care programs.
3. Data for FY12 through FY14 are not presented in the enrollment vs. spending analysis because no spending was reported on the 'Long Term Managed Care' line in the New York State Medicaid Global Cap Reports during those years. As a result, the analysis begins with FY15, the first year for which spending on that line was reported.
